\chapter{Mrs.\ Byass's Lodgings}

\textsc{``You} are Mr. Kirkwood?'' said his visitor civilly. ``My name
is Snowdon. I should be glad to speak a few words, with you, if you
could spare the time.''

Sidney's thoughts were instantly led into the right channel; he
identified the old man by his white hair and the cloak. The hat,
however, which had been described to him, was now exchanged for a soft
felt of a kind common enough; the guernsey, too, had been laid aside.
With ready good-will he invited Mr. Snowdon to enter.

There was not much in the room to distinguish it from the dwelling of
any orderly mechanic. A small bed occupied one side; a small table stood
before the window; the toilet apparatus was of course unconcealed; {}a
half-open cupboard allowed a glimpse of crockery, sundries, and a few
books. The walls, it is true, were otherwise ornamented than is usual;
engravings, chromo-lithographs, and some sketches of landscape in
pencil, were suspended wherever light fell, and the choice manifested in
this collection was nowise akin to that which ruled in Mrs. Peckover's
parlour, and probably in all the parlours of Tysoe Street. To select for
one's chamber a woodcut after Constable or Gainsborough is at all events
to give proof of a capacity for civilisation.

The visitor made a quick survey of these appearances; then he seated
himself on the chair Sidney offered. He was not entirely at his ease,
and looked up at the young man twice or thrice before he began to speak
again.

``Mr. Kirkwood, were you ever acquainted with my son, by name Joseph
Snowdon? ''

``No, I never knew him,'' was the reply. ``I have heard his name, and I
know where he once lived---not far from here.''

"You're wondering what has brought me {}to you. I have heard of you from
people a grandchild of mine is living with. I daresay it is the house
you mean,---in Clerkenwell Close."

``So you have found it!'' exclaimed Sidney with pleasure. ``I've been
looking about for you as I walked along the streets these last two or
three days.''

``Looking for me?'' said the other, astonished.

Sidney supplied the explanation, but without remarking on the
circumstances which made Jane so anxious to discover a possible friend.
Snowdon listened attentively, and at length, with a slight smile; he
seemed to find pleasure in the young man's way of expressing himself.
When silence ensued, he looked about absently for a moment; then,
meeting Sidney's eyes, said in a grave voice:

``That poor child is very ill.''

``Ill? I'm sorry to hear it.''

"The reason I've come to you, Mr. Kirkwood, is because she's called out
your name so often. They don't seem able to tell me {}how she came into
this state, but she's had a fright of some kind, or she's been living
very unhappily. She calls on your name, as if she wanted you to protect
her from harm. I didn't know what to think about it at first. I'm a
stranger to everybody---I may tell you I've been abroad for several
years---and they don't seem very ready to put trust in me; but I decided
at last that I'd come and speak to you. It's my grandchild, and perhaps
the only one of my family left; nobody can give me news of her father
since he went away four or five years ago. She came to herself this
morning for a little, but I'm afraid she couldn't understand what I
tried to tell her; then I mentioned your name, and I could see it did
her good at once. What I wish to ask of you is, would you come to her
bedside for a few minutes? She might know you, and I feel sure it would
be a kindness to her."

Sidney appeared to hesitate. It was not, of course, that he dreamt of
refusing, but he was busy revolving all he knew of Jane's life with the
Peckovers, and asking himself {}what it behoved him to tell, what to
withhold. Daily experience guarded him against the habit of gossip,
which is one of the innumerable curses of the uneducated (whether poor
or wealthy), and, notwithstanding the sympathy with which his visitor
inspired him, he quickly decided to maintain reserve until he understood
more of the situation.

``Yes, yes, I'll go with you at once,'' he made haste to reply, when he
perceived that his hesitancy was occasioning doubt and trouble. ``In
fact, I was just starting to go and see the Hewetts when you knocked at
the door. They're friends of mine,---living in Mrs. Peckover's house.
That's how I came to know Jane. I haven't been there for several days,
and when I last saw her, as I was saying, she seemed as well as usual.''

``I'm afraid that wasn't much to boast of,'' said Snowdon. ``She's a
poor, thin-looking child.''

Sidney was conscious that the old man did not give expression to all he
thought. This mutual exercise of tact seemed, however, to {}encourage a
good understanding between them rather than the reverse.

``You remain in the house?'' Kirkwood asked, as they went downstairs.

``I stay with her through the night. I didn't feel much confidence in
the doctor that was seeing her, so I made inquiries and found a better
man.''

When they reached the Close, the door was opened to them by Clem
Peckover. She glared haughtily at Sidney, but uttered no word. To
Kirkwood's surprise, they went up to the Hewetts' back-room. The
mattress that formerly lay upon the floor had been removed; the bed was
occupied by the sick girl, with whom at present Mrs. Peckover was
sitting. That benevolent person rose on seeing Sidney, and inclined her
head with stateliness.

``She's just fell asleep,'' was her whispered remark. ``I shouldn't say
myself as it was good to wake her up, but of course you know best.''

This was in keeping with the attitude Mrs. {}Peckover had adopted as
soon as she understood Snowdon's resolve to neglect no precaution on the
child's behalf. Her sour dignity was meant to express that she felt hurt
at the intervention of others where her affections were so nearly
concerned. Sidney could not help a certain fear when he saw this woman
installed as sick-nurse. It was of purpose that he caught her eye and
regarded her with a gravity she could scarcely fail to comprehend.

Jane awoke from her fitful slumber. She looked with but half-conscious
fearfulness at the figures darkening her view. Sidney moved so that his
face was in the light, and, bending near to her, asked if she recognised
him. A smile---slow-forming, but unmistakable at last---amply justified
what her grandfather had said. She made an effort to move her hand
towards him. Sidney responded to her wish, and again she smiled,
self-forgetfully, contentedly. Snowdon turned to Mrs. Peckover, and,
after a few words with regard to the {}treatment that was being pursued,
said that he would now relieve her; she lingered, hut shortly left the
room. Sidney, sitting by the bed, in a few minutes saw that Jane once
more slept, or appeared to do so. He whispered to Snowdon that he was
going to see his friends in the next room, and would look in again
before leaving.

His tap at the door was answered by Amy, who at once looked back and
said:

``Can Mr. Kirkwood come in, mother?''

``Yes; I want to see him,'' was the answer.

Mrs. Hewett was lying in bed: she looked, if possible, more wretchedly
ill than four days ago. On the floor were two mattresses, covered to
make beds for the children. The baby, held in its mother's aims, was
crying feebly.

``Why, I hoped you were getting much better by now,'' said Sidney.

Mrs. Hewett told him that she had been to the hospital on Saturday, and
seemed to have caught cold. A common enough occurrence; hours of waiting
in an out-patients' {}room frequently do more harm than the doctor's
advice can remedy. She explained that Mrs. Peckover had requested the
use of the other room.

``There's too many of us to be livin' an' sleepin' in this little
place,'' she said; ``but after all, it's a savin' of rent. It's a good
thing Clara isn't here.---An' you've heard as John's got work?''

He had found a job at length with a cabinetmaker; to-night he would
probably be working till ten or eleven o'clock. Good news so far. Then
Mrs. Hewett began to speak with curiosity of the old man who claimed
Jane as his grandchild. Sidney told her what had just happened.

``An' what did you say about the girl?'' she asked anxiously.

``I said as little as I could; I thought it wisest. Do you know what
made her ill?''

``It was that Clem as did it,'' Mrs. Hewett replied, subduing her voice.
And she related what had befallen after Sidney's last visit.

"Mrs. Peckover, she's that afraid the truth {}should get out. Of course
I don't want to make no bother, but I do feel that glad the poor thing's
got somebody to look after her at last. I never told you half the things
as used to go on. That Clem's no better than a wild-beast tiger; but
then what can you do? There's never any good comes out of makin' a
bother with other people's business, is there? Fancy him comin' to see
you! Mrs. Peckover's afraid of him, I can see that, though she pretends
she isn't goin' to stand him interferin'. What do you think about him,
Sidney? He's sent for a doctor out of Islington; wouldn't have nothin'
to say to the other. He must have plenty of money, don't you think? Mrs.
Peckover says he's goin' to pay the money owin' to her for Jane's keep.
As if the poor thing hadn't more than paid for her bits of meals an' her
bed in the kitchen! Do you think that woman 'ud ever have kept her if it
wasn't she could make her a servant with no wages? If Jane 'ud been a
boy, she'd a gone to the workhouse long ago. She's been that handy,
{}poor little mite! I've always done what I could for her; you know
that, Sidney. I do hope she'll get over it. If anything happens, mind my
word, there'll be a nice to do! Clara says she'll go to a magistrate an'
let it all out, if nobody else will. She hates the Peckovers, Clara
does."

``It won't come to that,'' said Sidney. ``I can see the old man'll take
her away as soon as possible. He may have a little money; he's just come
back from Australia. I like the look of him myself.''

He began to talk of other subjects; waxed wrath at the misery of this
housing to which the family had shrunk; urged a removal from the vile
den as soon as ever it could be managed. Sidney always lost control of
himself when he talked with the Hewetts of their difficulties; the
people were, from his point of view, so lacking in resource, so
stubbornly rooted in profitless habit. Over and over again he had
implored them to take a rational view of the case, to borrow a few
pounds of him, to make a new beginning on clean soil. {}It was like
contending with some hostile force of nature; he spent himself in vain.
As Hewett did not return, he at length took his leave, and went into the
back-room for a moment.

``She's asleep,'' said Snowdon, rising from the chair where he had been
sitting deep in thought. ``It's a good sign.''

Sidney just looked towards the bed, and nodded with satisfaction. The
old man gave him a warm pressure of the hand, and he departed. All the
way home, he thought with singular interest of the bare sick-room, of
the white-headed man watching through the night; the picture impressed
him in a way that could not be explained by its natural pathos merely;
it kept suggesting all sorts of fanciful ideas, due in a measure,
possibly, to Mrs. Hewett's speculations. For an hour he was so lost in
musing on the subject that he even rested from the misery of his
ceaseless thought of Clara.

He allowed three days to pass, then went to inquire about Jane's
progress. It had been {}satisfactory. Subsequent visits brought him to
terms of a certain intimacy with Snowdon. The latter mentioned at length
that he was looking for two rooms, suitable for himself and Jane. He
wished them to be in a decent house, somewhere in Clerkenwell, and the
rent was not to be more than a working-man could afford.

``You don't know of anything in your street?'' he asked diffidently.

Something in the tone struck Sidney. It half expressed a wish to live in
his neighbourhood if possible. He looked at his companion (they were
walking together), and was met in return with a glance of calm
friendliness; it gratified him, strengthened the feeling of respect and
attachment which had already grown out of this intercourse. In Tysoe
Street, however, no accommodation could be found. Sidney had another
project in his thoughts; pursuing it, he paid a visit the next evening
to certain acquaintances of his named Byass, who had a house in Hanover
Street, Islington, and let lodgings. Hanover Street {}lies to the north
of City Road; it is a quiet byway, of curving form, and consists of
dwellings only. Squalor is here kept at arm's length; compared with
regions close at hand, this and the contiguous streets have something of
a suburban aspect.

Three or four steps led up to the house-door. Sidney's knock summoned a
young, healthy-faced, comely woman, who evinced hearty pleasure on
seeing who her visitor was. She brought him at once into a parlour on
the ground-floor.

"Well, an' I was only this mornin' tellin' Sam to go an' look after you,
or write a note, or somethin'! Why can't you come round oftener? I've no
patience with you! You just sit at `ome an get humped, an' what's the
good o' that, I should like to know? I thought you'd took offence with
me, an so I told Sam. Do you want to know how baby is? Why don't you
ask, then, as you ought to do the first thing? He's a good deal better
than he deserves to be, young rascal,---all the trouble he gives me!
He's fast {}asleep, I'm glad to say, so you can't see him. Sam'll be
back in a few minutes; at least I expect him, but there's no knowin'
now-a-days when he can leave the warehouse. What's brought you to-night,
I wonder? You needn't tell me anything about the Upper Street business;
\emph{I} know all about \emph{that}!"

``Oh, do you? From Clara herself?''

``Yes. Don't talk to me about her! There! I'm sick an' tired of
her,---an so are you, I should think, if you've any sense left. Her an'
me can't get along, an' that's the truth. Why, when I met her on Sunday
afternoon, she was that patronisin' you'd have thought she'd got a place
in Windsor Castle. Would she come an' have a cup of tea? Oh, dear no!
Hadn't time! The Princess of Wales, I suppose, was waitin round the
corner!''

Having so relieved her mind, Mrs. Byass laughed with a genuine gaiety
which proved how little malice there was in her satire. Sidney could not
refuse a smile, but it was a gloomy one.

{}``I'm not sure you've done all you might have to keep her friends with
you,'' he said seriously, but with a good-natured look.

``There you go!'' exclaimed Mrs. Byass, throwing hack her head. "Of
course everybody must be in fault sooner than \emph{her}! She's an angel
is Miss Hewett! Poor dear! to think how shameful she's been used! Now I
do wonder how you've the face to say such things, Mr. Kirkwood! Why,
there's nobody else livin' would have been as patient with her as I
always was. I'm not bad-tempered, I will say that for myself, an' I've
put up with all sorts of things (me, a married woman), when any one else
would have boxed her ears an' told her she was a conceited minx. I used
to be fond of Clara; you know I did. But she's got beyond all bearin';
and if you wasn't just as foolish as men always are, you'd see her in
her true colours. Do shake yourself a bit, do! Oh, you silly, silly
man!"

Again she burst into ringing laughter, throwing herself backwards and
forwards, and {}at last covering her face with her hands. Sidney looked
annoyed, but the contagion of such spontaneous merriment in the end
brought another smile to his face. He moved his head in sign of giving
up the argument, and, as soon as there was silence, turned to the object
of his visit.

``I see you've still got the card in the window. I shouldn't wonder if I
could find you a lodger for those top-rooms.''

``And who's that? No children, mind.''

Sidney told her what he could of the old man. Of Jane he only said that
she had hitherto lived with the Hewetts' landlady, and was now going to
be removed by her grandfather, having just got through an illness. Dire
visions of infection at once assailed Mrs. Byass; impossible to admit
under the same roof with her baby a person who had just been ill. This
scruple was, however, overcome; the two rooms at the top of the
house---unfurnished---had been long vacant, owing to fastidiousness in
Mr. and Mrs. Byass, since their last lodger, after a fortnight of
{}continuous drunkenness, broke the windows, ripped the paper off the
walls, and ended by trying to set fire to the house. Sidney was
intrusted with an outline treaty, to be communicated to Mr. Snowdon.

This discussion was just concluded when Mr. Samuel Byass presented
himself,---a slender, large-headed young man, with very light hair,
cropped close upon the scalp, and a foolish face screwed into an
expression of facetiousness. He was employed in some clerkly capacity at
a wholesale stationer's in City Road. Having stepped into the room, he
removed a very brown silk hat and laid it on a chair, winking the while
at Sydney with his right eye; then he removed his overcoat, winking with
the left eye. Thus disembarrassed, he strode gravely to the fireplace,
took up the poker, held it in the manner of a weapon upright against his
shoulder, and exclaimed in a severe voice, ``Eyes right!'' Then
converting the poker into a sword, he drew near to Sydney and affected
to practise upon him the military cuts, his features {}distorted into
grotesque ferocity. Finally, assuming the attitude of a juggler, he made
an attempt to balance the poker perpendicularly upon his nose, until it
fell with a crash, just missing the ornaments on the mantelpiece. All
this time Mrs. Byass shrieked with laughter, with difficulty keeping her
chair.

``Oh, Sam,'' she panted forth, her handkerchief at her eyes, ``what a
fool you are! Do stop, or you'll kill me!''

Vastly gratified, Samuel advanced with ludicrous gestures towards the
visitor, held out his hand, and said with affected nasality, ``How do
you do, sir? It's some time since I had the pleasure of seeing you, sir.
I hope you have been pretty tolerable.''

"\emph{Isn't} he a fool, Mr. Kirkwood?" cried the delighted wife. ``Do
just give him a smack on the side of the head, to please me! Sam, go an
wash, an' we'll have supper. What do you mean by being so late
to-night?''

``Where's the infant?'' asked Mr. Byass, thrusting his hands into his
waistcoat pockets and peering about the room. "Bring forth {}the infant!
Let a fond parent look upon his child."

``Go an wash or I'll throw something at you. Baby's in bed, and mind,
you wake him if you dare!''

Sidney would have taken his leave, but found it impossible. Mrs. Byass
declared that if he would not stay to supper he should never enter the
house again.

``Let's make a night of it!'' cried Sam, standing in the doorway.
``Let's have three pots of six ale and a bottle of old Tom! Let us be
reckless!''

His wife caught up the pillow from the sofa and hurled it at him. Samuel
escaped just in time. The next moment his head was again thrust forward.

``Let's send to the High Street for three cold roast-fowls and a
beef-steak pie! Let's get custards and cheese-cakes and French pastry!
Let's have a pine-apple and preserved ginger! Who says, Go it for
once?''

Mrs. Byass caught up the poker and sprang {}after him. From the passage
came sounds of scuffling and screaming, and in the end of something
produced by the lips. Mrs. Byass then showed a very red face at the
door, and said:

"\emph{Isn't} he a fool? Just wait a minute while I get the table laid."

Supper was soon ready in the comfortable kitchen. A cold shoulder of
mutton, a piece of cheese, pickled beetroot, a seed-cake and raspberry
jam; such was the fare to which Bessie Byass invited her husband and her
guest. On a side-table were some open cardboard boxes containing
artificial flowers and leaves, for Bessie had now and then a little
``mounting'' to do for a shop in Upper Street, and in that way aided the
income of the family. She was in even better spirits than usual at the
prospect of letting her top-rooms. On hearing that piece of news,
Samuel, who had just come from the nearest public-house with a foaming
jug, executed a wild dance round the room and inadvertently knocked two
plates from the dresser. This accident made {}his wife wrathful, but
only for a moment; presently she was laughing as unrestrainedly as ever,
and bestowing upon the repentant young man her familiar flattery.

At eleven o'clock Sidney left them, and mused with smiles on his way
home. This was not exactly his ideal of domestic happiness, yet it was
better than the life led by the Hewetts---better than that of other
households with which he was acquainted---better far, it seemed to him,
than the aspirations which were threatening to lead poor Clara---who
knew whither? A temptation beset him to walk round into Upper Street and
pass Mrs. Tubbs's bar. He resisted it, knowing that the result would
only be a night of sleepless anger and misery.

The next day he again saw Snowdon, and spoke to him of Mrs. Byass's
rooms. The old man seemed at first indisposed to go so far, but when he
had seen the interior of the house and talked with the landlady, his
objections disappeared. Before another week had passed, the two rooms
were furnished {}in the simplest possible way, and Snowdon brought Jane
from Clerkenwell Close.

Kirkwood came by invitation as soon as the two were fairly established
in their home. He found Jane sitting by the fire in her grandfather's
room; a very little exertion still out-wearied her, and the strange
things that had come to pass had made her habitually silent. She looked
about her wonderingly, seemed unable to realise her position, was
painfully conscious of her new clothes, ever and again started as if in
fear.

``Well, what did I say that night?'' was Sidney's greeting ``Didn't I
tell you it would be all right soon?''

Jane made no answer in words, but looked at him timidly; and then a
smile came upon her face, an expression of joy that could not trust
itself, that seemed to her too boldly at variance with all she had yet
known of life.
