\part{}

\chapter{A Welcome Guest}

\textsc{The} bells of St. James's, Clerkenwell, ring melodies in
intervals of the pealing for service-time. One morning of spring their
music, like the rain that fell intermittently, was flung westwards by
the boisterous wind, away over Clerkenwell Close, until the notes failed
one by one, or were clashed out of existence by the clamour of a less
civilised steeple. Had the wind been under mortal control, it would
doubtless have blown thus violently and in this quarter in order that
the inhabitants of the House of Detention might derive no solace from
the melody. Yet I know not; just now the bells were {}playing ``There is
a happy land, far, far away,'' and that hymn makes too great a demand
upon the imagination to soothe amid instant miseries.

In Mrs. Peckover's kitchen the music was audible in bursts. Clem and her
mother, however, it neither summoned to prepare for church, nor lulled
into a mood of restful reverie. The two were sitting very close together
before the fire, and holding intimate converse; their voices kept a low
murmur, as if, though the door was shut, they felt it necessary to use
every precaution against being overheard. Three years have come and gone
since we saw these persons. On the elder, time has made little
impression; but Clem has developed noticeably. The girl is now in the
very prime of her ferocious beauty. She has grown taller and somewhat
stouter; her shoulders spread like those of a caryatid; the arm with
which she props her head is as strong as a carter's and magnificently
moulded. The head itself looks immense with its pile of glossy hair.
{}Reddened by the rays of the fire, her features had a splendid savagery
which seemed strangely at discord with the paltry surroundings amid
which she sat; her eyes just now were gleaming with a crafty and cruel
speculation which would have become those of a barbarian in ambush. I
wonder how it came about that her strain, after passing through the
basest conditions of modern life, had thus reverted to a type of
ancestral exuberance.

``If only he doesn't hear about the old man or the girl from somebody!''
said Mrs. Peckover. ``I've been afraid of it ever since he come into the
'ouse. There's so many people might tell him. You'll have to come round
him sharp, Clem.''

The mother was dressed as her kind are wont to be on Sunday
morning,---that is to say, not dressed at all, but hung about with
coarse garments, her hair in unbeautiful disarray. Clem, on the other
hand, seemed to have devoted much attention to her morning toilet; she
wore a dark dress trimmed with {}velveteen, and a metal ornament of
primitive taste gleamed amid her hair.

``There ain't no mistake?'' she asked, after a pause. ``You're jolly
sure of that?''

``Mistake? What a blessed fool you must be! Didn't they advertise in the
papers for him? Didn't the lawyers themselves say as it was somethink to
his advantage? Don't you say yourself as Jane says her grandfather's
often spoke about him and wished he could find him? How can it be a
mistake? If it was only Bill's letter we had to go on, you might talk;
but---there, don't be a ijiot!''

``If it turned out as he hadn't nothing,'' remarked Clem resolutely,
``I'd leave him, if I was married fifty times.''

Her mother uttered a contemptuous sound. At the same time she moved her
head as if listening; some one was, in fact, descending the stairs.

``Here he comes,'' she whispered. ``Get the eggs ready, an I'll make the
corffee.''

A tap at the door, then entered a tallish {}man of perhaps forty, though
he might be a year or two younger. His face was clean-shaven,
harsh-featured, unwholesome of complexion; its chief peculiarity was the
protuberance of the bone in front of each temple, which gave him a
curiously animal aspect. His lower lip hung and jutted forward; when he
smiled, as now in advancing to the fire, it slightly overlapped the one
above. His hair was very sparse; he looked, indeed, like one who has
received the tonsure. The movement of his limbs betokened excessive
indolence; he dragged his feet rather than walked. His attire was
equally suggestive; not only had it fallen into the last degree of
shabbiness, (having originally been such as is worn by a man above the
mechanic ranks), but it was patched with dirt of many kinds, and held
together by a most inadequate supply of buttons. At present he wore no
collar, and his waistcoat, half-open, exposed a red shirt.

``Why, you're all a-blowin' and a-growin' this morning. Miss Peckover,''
was his first observation, as he dropped heavily into a {}wooden
arm-chair. ``I shall begin to think that colour of yours ain't natural.
Dare you let me rub it with a handkerchief?''

``Course I dare,'' replied Clem, tossing her head. ``Don't be so
forward, Mr. Snowdon.''

``Forward? Not I. I'm behind time if anything. I hope I haven't kept you
from church.''

He chuckled at his double joke. Mother and daughter laughed
appreciatively.

``Will you take your eggs boiled or fried?'' inquired Mrs. Peckover.

``Going to give me eggs, are you? Well, I've no objection, I assure you.
And I think I'll have them fried, Mrs. Peckover. But, I say, you mustn't
be running up too big a bill. The Lord only knows when I shall get
anything to do, and it ain't very likely to be a thousand a year when it
does come.''

``Oh, that's all right,'' replied the landlady, as if sordid calculation
were a thing impossible to her. ``I can't say as you behaved quite
straightforward years ago, Mr. Snowdon, but I ain't one to make a row
about bygones, {}an' as you say you'll put it all straight as soon as
you can, well, I won't refuse to trust you once more.''

Mr. Snowdon lay back in the chair, his hands in his waistcoat pockets,
his legs outstretched upon the fender. He was smiling placidly, now at
the preparing breakfast, now at Clem. The latter he plainly regarded
with much admiration, and cared not to conceal it. When, in a few
minutes, it was announced to him that the meal was ready, he dragged his
chair up to the table and reseated himself with a sigh of satisfaction.
A dish of excellent ham, and eggs as nearly fresh as can be obtained in
Clerkenwell, invited him with appetising odour; a large cup of what is
known to the generality of English people as coffee steamed at his right
hand; slices of new bread lay ready cut upon a plate; a slab of the most
expensive substitute for butter caught his eye with yellow promise;
vinegar and mustard appealed to the refinements of his taste.

``I've got a couple more eggs, if you'd like {}them doin','' said Mrs.
Peckover, when she had watched the beginning of his attack upon the
viands.

``I think I shall manage pretty well with this supply,'' returned Mr.
Snowdon.

As he ate he kept silence, partly because it was his habit, partly in
consequence of the activity of his mind. He was, in fact, musing upon a
question which he found it very difficult to answer in any satisfactory
way. ``What's the meaning of all this?'' he asked himself, and not for
the first time. ``What makes them treat me in this fashion? A week ago I
came here to look up Mrs. Peckover, just because I'd run down to my last
penny, and I didn't know where to find a night's lodging. I'd got an
idea, too, that I should like to find out what had become of my child,
whom I left here nine or ten years ago; possibly she was still alive,
and might welcome the duty of supporting her parent. The chance was, to
be sure, that the girl had long since been in her grave, and that Mrs.
Peckover no longer lived in the old quarters; if I discovered the woman,
{}on the other hand, she was not very likely to give me an affectionate
reception, seeing that I found it inconvenient to keep sending her money
for Jane's keep in the old days. The queer thing is, that everything
turned out exactly the opposite of what I had expected. Mrs. Peckover
had rather a sour face at first, but after a little talk she began to
seem quite glad to see me. She put me into a room, undertook to board me
for a while,---till I find work, and I wonder when \emph{that}'ll
be?---and blest if this strapping daughter of hers doesn't seem to have
fallen in love with me from the first go off! As for my girl, I'm told
she was carried off by her grandfather, my old dad, three years ago, and
where they went nobody knows. Very puzzling all this. How on earth came
it that Mrs. Peckover kept the child so long, and didn't send her to the
workhouse? If I'm to believe \emph{her}, she took a motherly kindness
for the poor brat. But that won't exactly go down with J. J. Snowdon;
he's seen a bit too much in his knocking about the world. Still,---what
if I'm making a mistake about {}the old woman? There \emph{are} some
people do things of that sort; upon my soul, I've known people be kind
even to me, without a chance of being paid back! You may think you know
a man or a woman, and then all at once they'll go and do something you'd
have taken your davy couldn't possibly happen. I'd have sworn she was
nothing but a skinflint and a lying old witch. And so she may be; the
chances are there's some game going on that I can't see through. Make
inquiries? Why, so I have done, as far as I know how. I've only been
able to hit on one person who knows anything about the matter, and he
tells me it's true enough the girl was taken away about three years ago,
but he's no idea where she went to. Surely the old man must be dead by
now, though he was tough. Well, the fact of the matter is, I've got a
good berth, and I'm a precious sight too lazy to go on the private
detective job. Here's this girl Clem, the finest bit of flesh I've seen
for a long time; I've more than half a mind to see if she won't be fool
enough to marry {}me. I'm not a bad-looking fellow, that's the truth,
and she may have taken a real liking to me. Seems to me that I should
have come in for a comfortable thing in my old age; if I haven't a
daughter to provide for my needs, at all events I shall have a wife who
can be persuaded into doing so. When the old woman gets out of the way I
must have a little quiet talk with Clem.''

The opportunity he desired was not long in offering itself Having made
an excellent breakfast, he dragged his chair up to the fender again, and
reached a pipe from the mantelpiece, where he had left it last night.
Tobacco he carried loose in his waistcoat pocket; it came forth in the
form of yellowish dust, intermingled with all sorts of alien scraps.
When he had lit his pipe, he poised the chair on its hind-legs, clasped
his hands over his bald crown, and continued his musing with an air of
amiable calm. Smoke curled up from the corner of his loose lips, and
occasionally, removing his pipe for an instant, he spat skilfully
between the bars {}of the grate. Assured of his comfort, Mrs. Peckover
said she must go and look after certain domestic duties. Her daughter
had begun to clean some vegetables that would be cooked for dinner.

``How old may you be, Clem?'' Mr. Snowdon inquired genially, when they
had been alone together for a few minutes.

``What's that to you? Guess.''

``Why, let me see; you was not much more than a baby when I went away.
You'll be eighteen or nineteen, I suppose.''

``Yes, I'm nineteen---last sixth of February. Pity you come too late to
give me a birthday present, ain't it?''

``Ah! And who'd have thought you'd have grown up such a beauty! I say,
Clem, how many of the young chaps about here have been wanting to marry
you, eh?''

``A dozen or two, I dessay,'' Clem replied, shrugging her shoulders
scornfully. Mr. Snowdon laughed, and then spat into the fire.

``Tell me about some o' them, will you? {}Who is it you're keeping
company with now?''

``Who, indeed? Why, there isn't one I'd look at! Several of 'em's took
to drinking `cause I won't have nothing to do with 'em.''

This excited Mr. Snowdon's mirth in a high degree; he rolled on his
chair, and almost pitched backwards.

``I suppose you give one or other a bit of encouragement now and then,
just to make a fool of him, eh?''

``Course I do. There was Bob Hewett; he used to lodge here, but that was
after your time. I kep' him off an' on till he couldn't bear it no
longer; then he went an' married a common slut of a thing, just because
he thought it `ud make me mad. Ha, ha! I believe he'd give her poison
an' risk it any day, if only I promised to marry him afterwards. Then
there was a feller called Jeck Bartley. I set him an' Bob fightin' one
Bank-holiday,---you should a' seen 'em go at it! Jack went an' got
married a year ago to a girl called Suke Jollop; her mother forced
{}him. How I did laugh! Last Christmas Day they smashed up their 'ome
an' threw the bits out into the street. Jack got one of his eyes knocked
out,---I thought I should a' died o' laughin' when I saw him next
mornin'.''

The hearer became uproarious in merriment.

``Tell you what it is, Clem,'' he cried, ``you're something like a girl!
Darn me if I don't like you! I say, I wonder what my daughters grown up?
Like her mother, I suppose. You an' she was sort of sisters, wasn't
you?''

He observed her closely. Clem laughed and shrugged her shoulders.

``Queer sort o' sisters. She was a bit too quiet-like for me. There
never was no fun in her.''

``Aye, like her mother. And where did you say she went to with the old
man?''

``Where she went to?'' repeated Clem, regarding him steadily with her
big eyes, ``I never said nothing about it, `cause I didn't know.''

``Well, I shan't cry about her, and I don't {}suppose she misses me
much, wherever she is. All the same, Clem, I'm a domesticated sort of
man; you can see that, can't you? I shouldn't wonder if I marry again
one of these first days. Just tell me where to find a girl of the right
sort. I daresay you know heaps.''

``Dessay I do. What sort do you want?''

``Oh, a littlish girl,---yellow hair, you know,---one of them that look
as if they didn't weigh half-a-stone.''

``I'll throw this parsnip at you, Mr. Snowdon!''

``What's up now? You don't call yourself littlish, do you?''

Clem snapped the small end off the vegetable she was paring, and aimed
it at his head. He ducked just in time. Then there was an outburst of
laughter from both.

``Say, Clem, you haven't got a glass of beer in the house?''

``You'll have to wait till openin' time,'' replied the girl sourly,
going away to the far end of the room.

``Have I offended you, Clem?''

{}``Offended, indeed! As if I cared what you say!''

``Do you care what I think?''

``Not I!''

``That means you do. Say, Clem, just come here; I've something to tell
you.''

``You're a nuisance. Let me get on with my work, can't you?''

``No, I can't. You just come here. You'd better not give me the trouble
of fetching you!''

The girl obeyed him. Her cheeks were very hot, and the danger-signal was
flashing in her eyes. Ten minutes later she went upstairs, and had a
vivacious dialogue of whispers with Mrs. Peckover.
