\chapter{The Debt Repaid}

\textsc{She} rose early, in the murky cold of the winter morning. When,
at eight o'clock, she knocked as usual at her grandfather's door his
answer made her tremble.

``I shall be down in a few minutes, Jane I'll have breakfast with you.''

It was long since he had risen at this hour. His voice sounded less like
that of an old man, and, in spite of his calling her by her name, she
felt the tone to be severe. When he reached the parlour he did not offer
to take her hand, and she feared to approach him. She saw that his
features bore the mark of sleeplessness. Hers, poor girl! were yet more
woeful in their pallor. Through the meal he affected to occupy himself
with the book Miss Lant had {}the sight of which was intolerable to
Jane. And not for a full hour did he speak anything but casual words.
Jane had taken her sewing; unexpectedly he addressed her.

``Let's have a word or two together, Jane. I think we ought to, oughtn't
we?''

She forced herself to regard him.

``I think you meant what you said last night?''

``Grandfather, I will do whatever you bid me. I'll do it faithfully. I
was ungrateful. I feel ashamed to have spoken so.''

``That's nothing to do with it, Jane. You're not ungrateful; anything
but that. But I've had a night to think over your words. You couldn't
speak like that if you weren't driven to it by the strongest feeling you
ever knew or will know. I hadn't thought of it in that way; I hadn't
thought of you in that way.''

He began gently, but in the last words was a touch of reproof, almost of
scorn. He gazed at her from under his grey eyebrows, perhaps hoping to
elicit some resistance of her spirit, some sign of strength that would
help him to reconstruct his shattered ideal.

{}``Grandfather, I'll try with all my strength to be what you wish,---I
will!''

``And suppose the strength isn't sufficient, child?''

Even in her humility she could not but feel that this was unjust. Had
she ever boasted? Had she ever done more than promise tremblingly what
he demanded? But the fear was legitimate. A weak thing, all but
heartbroken, could she hope to tread firmly in any difficult path? She
hung her head, making no answer. He examined her, seeming to measure the
slightness of her frame. Sad, unutterably sad, was the deep breath he
drew as he turned his eyes away again.

``Do you feel well this morning, Jane?''

``Yes, grandfather.''

``Have you slept?''

``I couldn't. You were grieving about me. I hoped never to have
disappointed you.''

He fell into reverie. Was he thinking of that poor wife of his, dead
long, long ago, the well-meaning girl of whom he had expected impossible
things? A second time had he thus erred, no longer with the excuse of
{}experience and hot blood. That cry of Jane's had made its way to his
heart. An enthusiast, he was yet capable of seeing by the common light
of day, when his affections were deeply stirred. And in the night he had
pondered much over his son's behaviour. Was he being deceived in that
quarter also, and there intentionally? Did Joseph know this child better
than he had done, and calculate upon her weakuess? The shock, instead of
disabling him, had caused a revival of his strength. He could walk more
firmly this morning than at any time since his accident. His brain was
clear and active; he knew that it behoved him to reconsider all he had
been doing, and that quickly, ere it was too late. He must even forget
that aching of the heart until he had leisure to indulge it.

``You shan't disappoint me, my dear,'' he said gravely. ``It's my own
fault if I don't take your kindness as you mean it. I have to go out,
Jane, but I shall be back to dinner. Perhaps we'll talk again
afterwards.''

Of late, on the rare occasions of his leaving the house, he had always
told her where he was going, and for what purpose; Jane {}stood that
this confidence was at an end. When he was gone she found occupation for
a short time, but presently could only sit over the fire, nursing her
many griefs. She was no longer deemed worthy of confidence; worse than
that, she had no more faith in herself. If Sidney learnt what had
happened he could not even retain his respect for her. In this way she
thought of it, judging Kirkwood by the ideal standard, which fortunately
is so unlike human nature; taking it for granted---so oppressed was her
mind by the habit of dwelling on artificial motives---that he only liked
her because he had believed her strong in purpose, forgetting altogether
that his love had grown before he was aware that anything unusual was
required of her. She did remember, indeed, that it was only the depth of
her love for him which had caused her disgrace; but, even if he came to
understand that, it would not, she feared, weigh in her favour against
his judgment.

It was the natural result of the influences to which she had been
subjected. Her mind, overwrought by resolute contemplation of ideas
beyond its scope, her gentle nature bent {}beneath a burden of duty to
which it was unequal, and taught to consider with painful solemnity
those impulses of kindness which would otherwise have been merely the
simple joys of life, she had come to distrust every instinct which did
not subserve the supreme purpose. Even of Sidney's conduct she could not
reason in a natural way. Instinct would have bidden her reproach him,
though ever so gently; was it well done to draw away when he must have
known how she looked for his aid? Her artificial self urged, on the
other hand, that he had not acted thus without some gravely considered
motive. What it was she could not pretend to divine; her faith in his
nobleness overcame every perplexity. Of the persons constituting this
little group and playing their several parts, she alone had fallen
altogether below what was expected of her. As humble now as in the days
of her serfdom, Jane was incapable of revolting against the tyranny of
circumstances. Life had grown very hard for her again, but she believed
that this was to a great extent her own fault, the outcome of her own
unworthy weakness.

{}At Michael's return she did her best to betray no idle despondency.
Their midday meal was almost as silent as breakfast had been; his eyes
avoided her, and frequently he lost himself in thought. As he was rising
from the table Jane observed an unsteadiness in his movement; he shook
his head mechanically and leaned forward on both his hands, as if
feeling giddy. She approached him, but did not venture to speak.

``I'll go upstairs,'' he said, having sighed slightly.

``May I come and read to you, grandfather?''

``Not just now, Jane. Go out whilst it's a bit fine.''

He went from the room, still with an unsteady walk. Reaching his own
room, where there was a cheerful fire, he sat down, and remained for a
long time unoccupied, save with his reflections. This chamber had
scarcely changed in a detail of its arrangement since he first came to
inhabit it. There was the chair which Sidney always used, and that on
which Jane had sat since she was the silent, frail child of thirteen.
Here had his vision taken form, growing more definite with the growth of
his {}granddaughter, seeming to become at length a splendid reality.
What talk had been held here between Kirkwood and himself whilst Jane
listened! All gone into silence; gone, too, the hope it had encouraged.

He was weary after the morning's absence from home, and fell into a
light slumber. Dreams troubled him. First he found himself in Australia;
he heard again the sudden news of his son's death; the shock awoke him.
Another dozing fit, and he was a young man with a wife and children to
support; haunted with the fear of coming to want; harsh, unreasonable in
his exactions at home. Something like a large black coffin came into his
dream, and in dread of it he again returned to consciousness.

All night he had been thinking of the dark story of long ago,---his
wife's form motionless on the bed,---the bottle which told him what had
happened. Why must that memory revive to trouble his last days? Part of
his zeal for the great project had come of a feeling that he might thus
in some degree repair his former ill-doing; Jane would be a providence
to many hapless women whose burden was as heavy as {}his own wife's had
been. Must he abandon that solace? In any case he could bestow his money
for charitable purposes, but it would not be the same, it would not
effect what he had aimed at.

Late in the afternoon he drew from the inner pocket of his coat a long
envelope and took thence a folded paper. It was covered with clerkly
writing, which he perused several times. At length he tore the paper
slowly across the middle, again tore the fragments, and threw them on to
the fire. . . .

Jane obeyed her grandfather's word and went out for an hour. She wished
for news of Pennyloaf, who had been ill, and was now very near the time
of her confinement. At the door of the house in Merlin Place she was
surprised to encounter Bob Hewett, who stood in a lounging attitude; he
had never appeared to her so disreputable,---not that his clothes were
worse than usual, but his face and hands were dirty, and the former was
set in a hang-dog look.

``Is your wife upstairs, Mr. Hewett?'' Jane asked, when he had nodded
sullenly in reply to her greeting.

{}``Yes; and somebody else too as could have been dispensed with.
There's another mouth to feed.''

``No, there ain't,'' cried a woman's voice just behind him. Jane
recognised the speaker, a Mrs. GrifFm, who lived in the house and was
neighbourly to Pennyloaf.

``There ain't?'' inquired Bob, gruffly.

``The child's dead.''

``Thank goodness for that, any way!''

Mrs. Griffin explained to Jane that the birth had taken place twelve
hours ago. Pennyloaf was ``very low,'' but not in a state to cause
anxiety; perhaps it would be better for Jane to wait until tomorrow
before seeing her.

``She didn't say `thank goodness,' '' added the woman, with a scornful
glance at Bob, ``but I don't think she's over sorry as it's gone, an'
small blame to her. There's some people as doesn't care much what sort
o' times she has,---not meanin' \emph{you}, Miss, but them as had
\emph{ought} to care.''

Bob looked more disreputable than ever. His eyes were fixed on Jane, and
with such a singular expression that the latter, meeting {}their gaze,
felt startled, she did not know why. At the same moment he stepped down
from the threshold and walked away without speaking.

``I shouldn't care to have \emph{him} for a `usband,'' pursued Mrs.
Griffin. ``Of course he must go an' lose his work, just when his wife's
wantin' a few little extries, as you may say.''

``Lost his work?''

``Day `fore yes`day. I don't like him, an' I don't like his ways; he'll
be gettin' into trouble before long, you mind what I say. His family's a
queer lot, `cordin' to what they tell. Do you know them, Miss?''

``I used to, a long time ago.''

``You knew his sister---her as is come 'ome?''

``His sister?''

``Her as was a actress. Mrs. Bannister was tellin' me only last night;
she had it from Mrs. Horrocks, as heard from a friend of hers as lives
in the Farrin'don Buildin's, where the Hewetts lives too. They tell me
it was in the Sunday paper, though I don't remember nothing about it at
the time. It seems as how a woman threw vitrol over her an' burnt her
face so as there's no knowin' her, an' she goes about with a veil, an'
`cause she can't get her {}own livin' no more, of course she's come back
'ome, for all she ran away an' disgraced herself shameful.''

Jane gazed fixedly at the speaker, scarcely able to oather the sense of
what was said. ``Miss Hewett, you mean? Mr. Hewett's eldest daughter?''

``So I understand.''

``She has come home? When?''

``I can't just say; but a few weeks ago, I believe. They say it's nearly
two months since it was in the paper.''

``Does Mrs. Hewett know about it?''

``I can't say. She's never spoke to me as if she did. And, as I tell
you, I only heard yes'day myself. If you're a friend of theirs, p'r'aps
I hadn't oughtn't to a' mentioned it. It just come to my lips in the way
o' talkin'. Of course I don't know nothin' about the young woman myself;
it's only what you comes to 'ear in the way o' talkin', you know.''

This apology was doubtless produced by the listener's troubled
countenance. Jane asked no further question, but said she would come to
see Pennyloaf on the morrow, and so took her leave.

{}At ten o'clock next morning, just when Jane was preparing for her
visit to Merlin Place, so possessed with anxiety to ascertain if
Pennyloaf knew anything about Clara Hewett that all her troubles were
for the moment in the background, Bessie Byass came running upstairs
with a strange announcement. Sidney Kirkwood had called, and wished to
see Miss Snowdon in private for a few minutes.

``Something must have happened,'' said Jane, her heart standing still.

Bessie had a significant smile, but suppressed it when she noticed the
agitation into which her friend was fallen.

``Shall I ask him up into the front room?''

Michael was in his own chamber, which he had not left this morning. On
going to the parlour Jane found her visitor standing in expectancy. Yes,
something had happened; it needed but to look at him to be convinced of
that. And before a word was spoken Jane knew that his coming had
reference to Clara Hewett, knew it with the strangest certainty.

``I didn't go to work this morning,'' Sidney began, ``because I was very
anxious to see {}you---alone. I have something to speak about---to tell
you.''

``Let us sit down.''

Sidney waited till he met her look; she regarded him without
self-consciousness, without any effort to conceal her agitated interest.

``You see young Hewett and his wife sometimes. Have you heard from
either of them that Clara Hewett is living with her father again?''

``Not from them. A person in their house spoke about it yesterday. It
was the first I had heard.''

``Spoke of Miss Hewett? In a gossiping way, do you mean?''

``Yes.''

``Then you know what has happened to her?''

``If the woman told the truth.''

There was silence.

``Miss Snowdon''{{------}}

``Oh, I don't like you to speak so. You used to call me Jane.''

He looked at her in distress. She had spoken impulsively, but not with
the kind of emotion the words seem to imply. It was for his {}sake, not
for hers, that she broke that formal speech.

``You called me so when I was a child, Mr. Kirkwood,'' she continued,
smiling for all she was so pale. ``It sounds as if something had
altered. You're my oldest friend, and won't you always be so? Whatever
you're going to tell me, surely it doesn't prevent us from being
friends, just the same as always?''

He had not seen her in her weakness, the night before last. As little as
he could imagine that, was he able to estimate the strength with which
she now redeemed her womanly dignity. His face told her what he had to
disclose. No question now of proving herself superior to common
feelings; it was Sidney who made appeal to her, and her heart went forth
to grant him all he desired.

``Jane,---dear, good Jane,---you remember what I said to you in the
garden at Danbury,---that I had forgotten her. I thought it was true.
But you know what a terrible thing has befallen her. I should be less
than a man if I could say that she is nothing to me.''

``Have you spoken to her?''

``I have asked her to be my wife. Jane, if {}I had come to you
yesterday, before going to her, and had told you what I meant to do, and
explained all I felt, how the love of years ago had grown in me again,
wouldn't you have given me a friendly hand?''

``Just like I do now Do you think I have forgotten one night when she
stood by me and saved me from cruel treatment, and then nursed me when I
fell ill?''

Neither of them had the habit of making long speeches. They understood
each other,---very nearly; sufficiently, at all events, to make the bond
of sympathy between them stronger than ever. Jane was misled a little,
for she thought that here was the explanation of Sidney's withdrawing
his word to her grandfather; doubtless he heard of the calamity when it
happened. But on a more essential point she fell into no misconception.
Did Sidney desire that she should?

He held her hand until she gently drew it a way.

``You will go up and tell grandfather,'' she said, gravely; then added,
before he could speak, ``But I'll just see him first for a minute. He
hasn't been out of his room this mornino; yet. Please wait here.''

{}She left bim, and Sidney fell back on his chair, woebegone,
distracted.

Michael, brooding sorrowfully, at first paid no heed to Jane when she
entered his room. It was not loner since he had risen, and his simple
breakfast, scarcely touched, was still on the table.

``Grandfather, Mr. Kirkwood is here, and wishes to speak to you.''

He collected himself, and, regarding her, became aware that she was
strongly moved.

``Wishes to see me, Jane? Then I suppose he came to see you first?''

Prepared now for anything unexpected, feeling that the links between
himself and these young people were artificial, and that he could but
watch, as if from a distance, the course of their lives, his first
supposition was, that Sidney had again altered his mind. He spoke
coldly, and had little inclination for the interview.

``Yes,'' Jane replied, ``he came to see me, but only to tell me that he
is going to be married.''

His wrinkled face slowly gathered an expression of surprise.

``He will tell you who it is; he will explain. {}But I wanted to speak
to you first. Grandfather, I was afraid you might say something about
me. Will you---will you forget my foolishness? Will you think of me as
you did before? When he has spoken to you, you will understand why I am
content to put everything out of my mind, everything you and I talked
of. But I couldn't bear for him to know how I have disappointed you.
Will you let me be all I was to you before? Will you trust me again,
grandfather? You haven't spoken to him yet about me, have you?''

Michael shook his head.

``Then you will let it be as if nothing had happened?
Grandfather''{{------}}

She bent beside him and took his hand. Michael looked at her with a
light once more in his eyes.

``Tell him to come. He shall hear nothing from me, Jane.''

``And you will try to forget it?''

``I wish nothing better. Tell him to come here, my child. When he's gone
we'll talk together again.''

The interview did not last long, and Sidney {}left the house without
seeing Jane a second time.

She would have promised anything now. Seeing that life had but one path
of happiness for her, the path hopelessly closed, what did it matter by
which of the innumerable other ways she accomplished her sad journey?
For an instant, whilst Sidney was still speaking, she caught a gleam of
hope in renunciation itself, the kind of strength which idealism is fond
of attributing to noble natures. A gleam only, and deceptive; she knew
it too well after the day spent by her grandfather's side, encouraging,
at the expense of her heart's blood, all his revived faith in her. But
she would not again give way. The old man should reap fruit of her
gratitude, and Sidney should never suspect how nearly she had proved
herself unworthy of his high opinion.

She had dreamed her dream, and on awaking must be content to take up the
day's duties. Just in the same way, when she was a child at Mrs.
Peckover's, did not sleep often bring a vision of happiness, of freedom
from bitter tasks, and had she not to wake in the {}miserable mornings,
trembling lest she had lain too long? Her condition was greatly better
than then, so much better that it seemed wicked folly to lament because
one joy was not granted her.---Why, in the meantime she had forgotten
all about Pennyloaf. That visit must be paid the first thing this
morning.
