\chapter{The Treasury Unlocked}

\textsc{A Sunday} morning. In their parlour in Burton Crescent, Mr. and
Mrs. Joseph Snowdon were breakfasting. The sound of church bells---most
depressing of all sounds that mingle in the voice of London---intimated
that it was nearly eleven o'clock, but neither of our friends had in
view the attendance of public worship. Blended odours of bacon and
kippered herrings filled the room---indeed, the house, for several
breakfasts were in progress under the same roof. For a wonder, the
morning was fine, even sunny; a yellow patch glimmered on the worn
carpet, and the grime of the windowpanes was visible against an
unfamiliar sky. Joseph, incompletely dressed, had a Suuday paper propped
before him, and read whilst he ate. Clem, also in anything but
\emph{grande toilette}, was using a knife for the purpose of
{\protect\hypertarget{157}{}{}}conveying to her mouth the juice which
had exuded from crisp rashers. As usual, they had very little to say to
each other. Clem looked at her husband now and then, from under her
eyebrows, surreptitiously. After one of these glances she said, in a
tone which was not exactly hostile, but had a note of suspicion:

``I'd give something to know why he's going to marry Clara Hewett.''

``Not the first time you've made that remark,'' returned Joseph, without
looking up from his paper.

``I suppose I can speak?''

``Oh yes. But I'd try to do so in a more lady-like way.''

Clem flashed at him a gleam of hatred. He had become fond lately of
drawing attention to her defects of breeding. Clem certainly did not
keep up with his own progress in the matter of external refinement; his
comments had given her a sense of inferiority, which irritated her
solely as meaning that she was not his equal in craft. She let a minute
or two pass, then returned to the subject.

``There's something at the bottom of it; I
{\protect\hypertarget{158}{}{}}know that. Of course you know more about
it than you pretend.''

Joseph leaned back in his chair and regarded her with a smile of the
loftiest scorn.

``It never occurs to you to explain it in the simplest way, of course.
If ever you hear of a marriage, the first thing you ask yourself is:
What has he or she to gain by it? Natural enough---in you. Now do you
really suppose that all marriages come about in the way that
\emph{yours} did---on your side, I mean?''

Clem was far too dull-witted to be capable of quick retort. She merely
replied:

``I don't know what you're talking about.''

``Of course not. But let me assure you that people sometimes think of
other things besides making profit when they get married. It's a pity
that you always show yourself so coarse-minded.''

Joseph was quite serious in administering this rebuke. He really felt
himself justified in holding the tone of moral superiority. The same
phenomenon has often been remarked in persons conscious that their
affairs are prospering, and whose temptations to paltry meanness are on
that account less frequent.

{\protect\hypertarget{159}{}{}}``And what about yourself?'' asked his
wife, having found her retort at length. ``Why did you want to marry
\emph{me}, I'd like to know?''

``Why? You are getting too modest. How could I live in the same house
with such a good-looking and sweet-tempered and well-behaved''{{------}}

``Oh, shut up!'' she exclaimed, in a voice such as one hears at the
street-corner. " It was just because you thought we was goin' to be
fools enough to keep you in idleness. Who was the fool, after all?''

Joseph smiled, and returned to his newspaper. In satisfaction at having
reduced him to silence, Clem laughed aloud and clattered with the knife
on her plate. As she was doing so there came a knock at the door.

``A gentleman wants to know if you're in, sir,'' said the house-thrall,
showing a smeary face. ``Mr. Byass is the name.''

``Mr. Byass? I'll go down and see him.''

Clem's face became alive with suspicion. In spite of her careless attire
she intercepted Joseph, and bade the servant ask Mr. Byass to come
upstairs. ``How can you go down without a collar?'' she said to her
husband.

{\protect\hypertarget{160}{}{}}He understood, and was somewhat uneasy,
but made no resistance. Mr. Byass presented himself. He had a very long
face, and obviously brought news of grave import. Joseph shook hands
with him.

``You don't know my wife, I think. Mr. Byass, Clem. Nothing wrong, I
hope?''

Samuel, having made his best City bow, swung back from his toes to his
heels, and stood looking down into his hat. ``I'm sorry to say,'' he
began, with extreme gravity, ``that Mr. Snowdon is rather ill---in fact,
very ill. Miss Jane asked me to come as sharp as I could.''

``Ill? In what way?''

``I'm afraid it's a stroke, or something in that line. He fell down
without a word of warning, just before ten o'clock. He's lying
insensible.''

``I'll come at once,'' said Joseph. ``They've got a doctor, I hope?''

``Yes; the doctor had been summoned instantly.''

``I'll go with you,'' said Clem, in a tone of decision.

``No, no; what's the good? You'll only be in the way.''

{\protect\hypertarget{161}{}{}}``No, I shan't. If he's as bad as all
that, I shall come.''

Both withdrew to prepare themselves. Mr. Byass, who was very nervous and
perspiring freely, began to walk round and round the table, inspecting
closely, in complete absence of mind, the objects that lay on it.

``We'll have a cab,'' cried Joseph, as he came forth equipped. ``Poor
Jane's in a sad state, I'm afraid, eh?''

In a few minutes they were driving up Pentonville Road. Clem scarcely
ever removed her eye from Joseph's face; the latter held his lips close
together and kept his brows wrinkled. Few words passed during the drive.
At the door of the house appeared Bessie, much agitated. All turned into
the parlour on the ground floor and spoke together for a few minutes.
Michael had been laid on his bed; at present Jane only was with him, but
the doctor would return shortly.

``Will you tell her I'm here? '' said Joseph to Mrs. Byass. ``I'll see
her in the sitting-room.''

He went up and waited. Throughout the house prevailed that unnatural,
nerve-distressing quietude which tells the presence of
{\protect\hypertarget{162}{}{}}calamity. The church bells had ceased
ringing, and Sunday's silence in the street enhanced the effect of
blankness and alarming expectancy. Joseph could not keep still; he
strained his ears in attention to any slight sound that might come from
the floor above, and his heart beat painfully when at length the door
opened.

Jane fixed her eyes on him and came silently forward.

``Does he show any signs of coming round?'' her father inquired.

``No. He hasn't once moved.''

She spoke only just above a whisper. The shock kept her still trembling
and her face bloodless.

``Tell me how it happened, Jane.''

``He'd just got up. I'd taken him his breakfast, and we were talking.
All at once he began to turn round, and then he fell down---before I
could reach him.''

``I'll go upstairs, shall I?''

Jane could not overcome her fear; at the door of the bedroom she drew
back, involuntarily, that her father might enter before her. When she
forced herself to follow, the first
{\protect\hypertarget{163}{}{}}glimpse of the motionless form shook her
from head to foot. The thought of death was dreadful to her, and death
seemed to lurk invisibly in this quiet room. The pale sunlight affected
her as a mockery of hope.

``You won't go away again, father?'' she whispered.

He shook his head.

In the meantime Bessie and Clem were conversing. On the single previous
occasion of Clem's visit to the house they had not met. They examined
each other's looks with curiosity. Clem wished it were possible to get
at the secrets of which Mrs. Byass was doubtless in possession; Bessie
on her side was reserved, circumspect.

``Will he get over it?'' the former inquired, with native brutality.

``I'm sure I don't know; I hope he may.''

The medical man arrived, and when he came downstairs again Joseph
accompanied him. Clem, when she found that nothing definite could be
learned, and that her husband had no intention of leaving, expressed her
wish to walk round to Clerkenwell Close and see her mother. Joseph
approved.

{\protect\hypertarget{164}{}{}}``You'd better have dinner there,'' he
said to her privately. ``We can't both of us come down on the Byasses.''

She nodded, and with a parting glance of hostile suspicion set forth.
When she had crossed City Road, Clem's foot was on her native soil; she
bore herself with conscious importance, hoping to meet some acquaintance
who would be impressed by her attire and demeanour. Nothing of the kind
happened, however. It was the dead hour of Sunday morning, midway in
service-time, and long before the opening of public-houses. In the
neighbourhood of those places of refreshment were occasionally found
small groups of men and boys, standing with their hands in their
pockets, dispirited, seldom caring even to smoke; they kicked their
heels against the kerbstone and sighed for one o'clock. Clem went by
them with a haughty balance of her head.

As she entered by the open front door and began to descend the kitchen
steps, familiar sounds were audible. Mrs. Peckover's voice was raised in
dispute with some one; it proved to be a quarrel with a female lodger
respecting {\protect\hypertarget{165}{}{}}the sum of
threepence-farthing, alleged by the landlady to be owing on some account
or other. The two women had already reached the point of calling each
other liar and thief. Clem, having no acquaintance with the lodger,
walked into the kitchen with an air of contemptuous indifference. The
quarrel continued for another ten minutes,---if the head of either had
been suddenly cut off it would assuredly have gone on railing for an
appreciable time,---and Clem waited, sitting before the fire. At last
the lodger had departed, and the last note of her virulence died away.

``And what do \emph{you} want? asked Mrs. Peckover, turning sharply upon
her daughter.

``I suppose I can come to see you, can't I?''

``Come to see me! Likely! When did you come last? You're a ungrateful
beast, that's what you are!''

``All right. Go a'ead! Anything else you'd like to call me?''

Mrs. Peckover was hurt by the completeness with which Clem had
established her independence. To do the woman justice, she had been
actuated, in her design of capturing Joseph Snowdon, at least as much by
a wish {\protect\hypertarget{166}{}{}}to establish her daughter
satisfactorily as by the ever- wakeful instinct which bade her seize
whenever gain lay near her clutches. Clem was proving disloyal, had
grown secretive. Mrs. Peckover did not look for any direct profit worth
speaking of from the marriage she had brought about, but she did desire
the joy of continuing to plot against Joseph with his wife. Moreover,
she knew that Clem was a bungler, altogether lacking in astuteness, and
her soul was pained by the thought of chances being missed. Her
encounter with the lodger had wrought her up to the point at which she
could discuss matters with Clem frankly. The two abused each other for a
while, but Clem really desired to communicate her news, so that calmer
dialogue presently ensued.

``Old Snowdon's had a stroke, if you'd like to know, and it's my belief
he won't get over it.''

``Your belief! And what's your belief worth? Had a stroke, has he? Who
told you?''

``I've just come from the `ouse. Jo's stoppin there.''

They discussed the situation in all its
{\protect\hypertarget{167}{}{}}aspects, but Mrs. Peckover gave it
clearly to be understood that, from her point of view, ``the game was
spoilt.'' As long as Joseph continued living under her roof she could in
a measure direct the course of events; Clem had chosen to abet him in
his desire for removal, and if ill came of it she had only herself to
blame.

``I can look out for myself,'' said Clem.

``Can you? I'm glad to hear it.''

And Mrs. Peckover sniffed the air, scornfully. The affectionate pair
dined together, each imbibing a pint and a half of ``mild and bitter,''
and Clem returned to Hanover Street. From Joseph she could derive no
information as to the state of the patient.

``If you will stay here, where you can do no good,'' he said, ``sit down
and keep quiet.''

``Certainly I shall stay,'' said his wife, ``because I know you want to
get rid of me.''

Joseph left her in the sitting-room, and went upstairs again to keep his
daughter company. Jane would not leave the bedside. To enter the room,
after an interval elsewhere, wrung her feelings too painfully; better to
keep her eyes fixed on the unmoving form, to overcome the dread by
facing it.

{\protect\hypertarget{168}{}{}}She and her father seldom exchanged a
word. The latter was experiencing human emotion, but at the same time he
had no little anxiety regarding his material interests. It was ten days
since he had learnt that there was no longer the least fear of a
marriage between Jane and Sidney, seeing that Kirkwood was going to
marry some one else,---a piece of news which greatly astonished him, and
confirmed him in his judgment that he had been on the wrong tack in
judging Kirkwood's character. At the same time he had been privily
informed by Scawthorne of an event which had ever since kept him very
uneasy,---Michael's withdrawal of his will from the hands of the
solicitors. With what purpose this had been done Scawthorne could not
conjecture; Mr. Percival had made no comment in his hearing. In all
likelihood the will was now in this very room. Joseph surveyed every
object again and again. He wondered whether Jane knew anything of the
matter, but not all his cynicism could persuade him that at the present
time her thoughts were taking the same direction as his own.

The day waned. Its sombre close was
{\protect\hypertarget{169}{}{}}unspeakably mournful in this haunted
chamber. Jane could not bear it; she hid her face and wept.

When the doctor came again, at six o'clock, he whispered to Joseph that
the end was nearer than he had anticipated. Near, indeed; less than ten
minutes after the warning had been given Michael ceased to breathe.

Jane knelt by the bed, convulsed with grief, unable to hear the words
her father addressed to her. He sat for five minutes, then again spoke.
She rose and replied.

``Will you come with us, Jane, or would you rather stay with Mrs.
Byass?''

``I will stay, please, father.''

He hesitated, but the thought that rose was even for him too io:noble to
be entertained.

``As you please, my dear. Of course no one must enter your rooms but
Mrs. Byass. I must go now, but I shall look in again to-night.''

``Yes, father.''

She spoke mechanically. He had to lead her from the room, and, on
quitting the house, left her all but unconscious in Bessie's arms.
