\chapter{The Heir}

\textsc{``And} you mean to say,'' cried Clem, when she was in the cab
with her husband speeding back to Burton Crescent---``you mean to say as
you've left them people to do what they like?''

``I suppose I know my own business,'' replied Joseph, wishing to convey
the very impression which in fact he did,---that he had the will in his
pocket. On reaching home he sat down at once and penned a letter to
Messrs. Percival \& Peel, formally apprising them of what had happened.
Clem sat by and watched him. Having sealed the envelope, he remarked:

``Pm going out for a couple of hours.''

``Then I shall go with you.''

``You'll do nothing of the kind. Why, what do you mean, you great,
gaping fool?''

The agitation of his nerves made him break
{\protect\hypertarget{171}{}{}}into unaccustomed violence. ``Do you
suppose you're going to follow me everywhere for the next week? Are you
afraid I shall run away? If I mean to do so, do you think you can stop
me? You'll just wait here till I come back, which will be before ten
o'clock. Do you hear?''

She looked at him fiercely, but his energy was too much for her, and
perforce she let him go. As soon as he had left the house, she too sat
down and indited a letter. It ran thus:

``\textsc{Dear Mother},---The old feller has gawn of it apened at jest
after six o'clock if you want to now I shall come and sea you at ten
e'clock to-morow moning and I beleve hes got the will but hes a beest
and theers a game up you may take your hothe so I remain C. S.''

This document she took to the nearest pillar-post, then returned and sat
brooding.

By the first hansom available Joseph was driven right across London to a
certain dull street in Chelsea. Before dismissing the vehicle he knocked
at the door of a lodging-house and made inquiry for Mr. Scawthorne. To
his surprise and satisfaction, Mr. {\protect\hypertarget{172}{}{}}thorne
happened to be at home; so the cabman was paid, and Joseph went up to
the second floor. In his shabby little room Scawthorne sat smoking and
reading. It was a season of impecuniosity with him, and his mood was
anything but cheerful. He did not rise when his visitor entered.

``Well now, what do you think brings me here?'' exclaimed Joseph, when
he had carefully closed the door.

``Hanged if I know, but it doesn't seem to be particularly bad news.''

Indeed, Joseph had overcome his sensibilities by this time, and his
aspect was one of joyous excitement. Seeing on the table a bottle of
sherry, loosely corked, he pointed to it.

``If you don't mind, Scaw. I'm a bit upset, a bit flurried. Got another
wine-glass?''

From the cupboard Scawthorne produced one, and bade the visitor help
himself His face began to express curiosity. Joseph tilted the draught
down his throat and showed satisfaction.

``That does me good. I've had a troublesome day. It ain't often my
feelings are tried.''

{\protect\hypertarget{173}{}{}}``Well, what is it?''

``My boy, we are all mortal. I dare say you've heard that observation
before; can you apply it to any particular case?''

Scawthorne was startled; he delayed a moment before speaking.

``You don't mean to say''{{------}}

``Exactly. Died a couple of hours ago, after lying insensible all day,
poor old man! I've just written your people a formal announcement. Now,
what do you think of that? If you don't mind, old fellow.''

He filled himself another glass, and tilted it off as before. Scawthorne
had dropped his eyes to the ground, and stood in meditation.

``Now, what about the will?'' pursued Joseph.

``You haven't looked for it?'' questioned his friend, with an odd look.

``Thought it more decent to wait a few hours. The girl was about, you
see, and what's more, my wife was. But have you heard anything since I
saw you?''

``Why, yes. A trifle.''

``Out with it! What are you grinning about? Don't keep me on hot
coals.''

``Well, it's amusing, and that's the fact.
{\protect\hypertarget{174}{}{}}Take another glass of sherry; you'll need
support.''

``Oh, I'm prepared for the worst. He's cut me out altogether, eh? That
comes of me meddling with the girl's affairs---damnation! When there
wasn't the least need, either.''

``A bad job. The fact is, Percival had a letter from him at midday
yesterday. The senior had left the office; young Percival opened the
letter, and spoke to me about it. Now, prepare yourself. The letter said
that he had destroyed his former will, and would come to the office on
Monday---that's \emph{to-morrow}---to give instructions for a new one.''

Joseph stood and stared.

``To-morrow? Why, then, there's \emph{no will at all?}''

``An admirable deduction. I congratulate you on your logic.''

Snowdon flung up his arms wildly, then began to leap about the room.

``Try another glass,'' said Scawthorne. ``There's still a bottle in the
cupboard; don't be afraid.''

``And you mean to tell me it's all mine?''

``The wine? You're very welcome.''

{\protect\hypertarget{175}{}{}}``Wine be damned! The money, my boy, the
money! Scawthorne, I'm not a mean chap. As sure as you and me stand
here, you shall have---you shall have a hundred pounds! I mean it; dash
me, I mean it! You've been devilish useful to me; and what's more---I
haven't done with you yet. Do you twig, old boy?''

``You mean that a confidential agent in England, unsuspected, may be
needed?''

``Shouldn't wonder if I do.''

``Can't be managed under double the money, my good sir,'' observed
Scawthorne, with unmistakable seriousness. ``Worth your while, I promise
you. Have another glass. Fair commission. Think it over.''

``Look here! I shall have to make the girl an allowance.''

``There's the filter- works. Don't be stingey.''

Joseph was growing very red in the face. He drank glass after glass; he
flung his arms about; he capered.

``Damn me if you shall call me that, Scaw! Two hundred it shall be. But
what was the old cove up to? Why did he destroy the other will? What
would the new one have been?''

{\protect\hypertarget{176}{}{}}``Can't answer either question, but it's
probably as well for you that \emph{to-morrow} never comes.''

``Now just see how things turn out!'' went on the other, in the joy of
his heart. ``All the thought and the trouble that I've gone through this
last year, when I might have taken it easy and waited for chance to make
me rich! Look at Kirkwood's business. There was you and me knocking our
heads together and raising lumps on them, as you may say, to find out a
plan of keeping him and Jane apart, when all the while we'd nothing to
do but to look on and wait, if only we'd known. Now this is what I call
the working of Providence, Scawthorne. Who's going to say, after this,
that things ain't as they should be? Everything's for the best, my boy;
I see that clearly enough.''

``Decidedly,'' assented Scawthorne, with a smile. ``The honest man is
always rewarded in the long-run. And that reminds me; I too have had a
stroke of luck.''

He went on to relate that his position in the office of Percival \& Peel
was now nominally that of an articled clerk, and that in three
{\protect\hypertarget{177}{}{}}years' time, if all went well, lie would
be received in the firm as junior partner.

``There's only one little project I am sorry to give up, in connection
with your affairs, Snowdon. If it had happened that your daughter had
inherited the money, why shouldn't I have had the honour of becominoo
your son-in-law?''

Joseph stared, then burst into hearty laughter.

``I tell you what,'' he said, recovering himself, ``why should you give
up that idea? She's as good a girl as you'll ever come across, I can
tell you that, my boy. There's better-looking, but you won't find many
as modest and good-hearted. Just make her acquaintance, and tell me if
I've deceived you. And look here, Scawthorne; by George, I'll make a
bargain with you! You say you'll be a partner in three years. Marry Jane
when that day comes, and I'll give you a thousand for a wedding present.
I mean it! What's more, I'll make my will on your marriage- day and
leave everything I've got to you and her. There now!''

``What makes you so benevolent all at once?'' inquired Scawthorne,
blandly.

{\protect\hypertarget{178}{}{}}``Do you think I've got no fatherly
feeling, man? Why, if it wasn't for my wife, I'd ask nothing; better
than to settle down with Jane to keep house for me. She's a good girl, I
tell you, and I wish her happiness.''

``And do you think I'm exactly the man to make her a model husband?''

``I don't see why not---now you're going to be a partner in a good
business. Don't you think I'm ten times as honest a man to-day as I was
yesterday? Poor devils can't afford to be what they'd wish, in the way
of honesty and decent living.''

``True enough, for once,'' remarked the other, without irony.

``You think it over, Scaw. I'm a man of my word. You shall have your
money as soon as things are straight: and if you can bring about that
affair, I'll do all I said,---so there's my hand on it. Say the word,
and I'll make you acquainted with her before---before I take that little
trip you know of, just for my health.''

``We'll speak of it again.''

Thereupon they parted. In the course of the following day Scawthorne's
report received {\protect\hypertarget{179}{}{}}official confirmation.
Joseph pondered deeply with himself whether he should tell his wife the
truth or not; there were arguments for both courses. By Tuesday morning
he had decided for the truth; that would give more piquancy to a
pleasant little jest he had in mind. At breakfast he informed her, as if
casually, and it amused him to see that she did not believe him.

``You'll be anxious to tell your mother. Go and spend the day with her,
but be back by five o'clock; then we'll talk things over. I have
business with the lawyers again.''

Clem repaired to the Close. Late in the afternoon she and her husband
again met at home, and by this time Joseph's elation had convinced her
that he was telling the truth. Never had he been in such a suave humour;
he seemed to wish to make up for his late severities. Seating himself
near her, he began pleasantly:

``Well, things might have been worse, eh?''

``I s'pose they might.''

``I haven't spoken to Jane yet. Time enough after the funeral. What
shall we do for the poor girl, eh?''

{\protect\hypertarget{180}{}{}}``How do I know?''

``You won't grudge her a couple of pounds a week, or so, just to enable
her to live with the Byasses, as she has been doing?''

``I s'pose the money's your own to do what you like with.''

``Very kind of you to say so, my dear. But we're well-to-do people now,
and we must be polite to each other. Where shall we take a house, Clem?
Would you like to be a bit out of town? There's very nice places within
easy reach of King's Cross, you know, on the Great Northern. A man I
know lives at Potter's Bar, and finds it very pleasant; good air. Of
course I must be within easy reach of business.''

She kept drawing her nails over a fold in her dress, making a scratchy
sound. ``It happened just at the right time,'' he continued. ``The
business wants a little more capital put into it. I tell you what it is,
Clem; in a year or two we shall be coining money, old girl.''

``Shall you?''

``Right enough. There's just one thing I'm a little anxious about; you
won't mind me {\protect\hypertarget{181}{}{}}mentioning it? Do you think
your mother'11 expect us to do anything for her?''

Clem regarded him with cautious scrutiny. He was acting well, and her
profound distrust began to be mingled with irritating uncertainty.

``What can she expect? If she does, she'll have to be disappointed,
that's all.''

``I don't want to seem mean, you know. But then she isn't so badly off
herself, is she?''

``I know nothing about it. You'd better ask her.'' And Clem grinned.
Thereupon Joseph struck a facetious note, and for half-an-hour made
himself very agreeable. Now for the first time, he said, could he feel
really settled; life was smooth before him. They would have a
comfortable home, the kind of place to which he could invite his
friends; one or two excellent fellows he knew would bring their wives,
and so Clem would have more society.

``Suppose you learn the piano, old girl? It wouldn't be amiss. By the
bye, I hope they'll turn you out some creditable mourning. You'll have
to find a West End dressmaker. She listened, and from time to time
smiled {ambiguously {.~.~.~.}} {\protect\hypertarget{182}{}{}}At noon of
the next clay Clem was walking on that part of the Thames Embankment
which is between Waterloo Bridge and the Temple Pier. It was a mild
morning, misty, but illuminated now and then with rays of sunlight,
which gleamed dully upon the river and gave a yellowness to remote
objects. At the distance of a dozen paces walked Bob Hewett; the two had
had a difference in their conversation, and for some minutes kept thus
apart, looking sullenly at the ground. Clem turned aside, and leaned her
arms on the parapet. Presently her companion drew near and leaned in the
same manner.

``What is it you want me to do?'' he asked huskily. ``Just speak plain,
can't you?''

``If you cant understand,---if you \emph{won't}, that is,---it's no good
speakin' plainer.''

``You said the other night as you didn't care about his money. If you
think he means hookin' it, let him go, and good riddance.''

``That's a fool's way of talkin'. I'm not goin' to lose it all, if I can
help it. There's a way of stoppin' him, and of gettin' the money too.''

They both stared down at the water; it
{\protect\hypertarget{183}{}{}}was full tide, and the muddy surface
looked almost solid.

``You wouldn't get it all,'' were Bob's next words. ``I've been asking
about that.''

``You have? Who did you ask?''

``Oh, a feller you don't know. You'd only have a third part of it, and
the girl `ud get the rest.''

``What do you call a third part?''

So complete was her stupidity, that B0I) had to make a laborious
explanation of this mathematical term. She could have understood what
was meant by a half or a quarter, but the unfamiliar ``third'' conveyed
no distinct meaning.

``I don't care,'' she said at length. " That `ud be enough.''

``Clem,---you'd better leave this job alone. You'd better, I warn you.''

``I shan't.''

Another long silence. A steamboat drew up to the Temple Pier, and a
yellow shaft of sunlight fell softly upon its track in the water.

``What do you want me to do?'' Bob recommenced. ``\emph{How?}''

Their eyes met, and in the woman's gaze he
{\protect\hypertarget{184}{}{}}found a horrible fascination, a devilish
allurement to that which his soul shrank from. She lowered her voice.

``There's lots of ways. It `ud be easy to make it seem as somebody did
it just to rob him. He's always out late at night.''

His face was much the colour of the muddy water yellowed by that shaft
of sunlight. His lips quivered. ``I dursn't, Clem. I tell you plain, I
dursn't.''

``Coward!'' she snarled at him, savagely.

``Coward! All right, Mr. Bob. You go your way, and I'll go mine.''

``Listen here, Clem,'' he gasped out, laying his hand on her arm. ``I'll
think about it. I won't say no. Give me a day to think about it.''

``Oh, we know what your thinkin' means.''

They talked for some time longer, and before they parted Bob had given a
promise to do more than think. The long, slouching strides with which he
went up from the embankment to the Strand gave him the appearance of a
man partly overcome with drink. For hours he walked about the City, in
complete oblivion of {\protect\hypertarget{185}{}{}}everything external.
Only when the lights began to shine from shop-windows did he consciously
turn to his own district. It was raining now. The splashes of cool
moisture made him aware how feverishly hot his face was.

When he got among the familiar streets he went slinkingly, hurrying
round corners, avoidinor glances. Almost at a run he turned into Merlin
Place, and he burst into his room as though he were pursued.

Pennyloaf had now but one child to look after, a girl of two years, a
feeble thing. Her own state was wretched; professedly recovered from
illness, she felt so weak, so lowspirited, that the greater part of her
day was spent in crying. The least exertion was too much for her; but
for frequent visits from Jane Snowdon she must have perished for very
lack of wholesome food. She was crying when startled by her husband's
entrance, and though she did her best to hide the signs of it, Bob saw.

``When are you going to stop that?'' he shouted.

She shrank away, looking at him with fear in her red eyes.

{\protect\hypertarget{186}{}{}}``Stop your snivelling, and get me some
tea!''

It was only of late that Pennyloaf had come to regard him with fear. His
old indifference and occasional brutality of language had made her life
a misery, but she had never looked for his return home with anything but
anxious longing. Now the anticipation was mingled with dread. He not
only had no care for her, not only showed that he felt her a burden upon
him; his disposition now was one of hatred, and the kind of hatred which
sooner or later breaks out in ferocity. Bob would not have come to this
pass---at all events not so soon---if he had been left to the dictates
of his own nature; he was infected by the savagery of the woman who had
taken possession of him. Her lust of cruelty crept upon him like a
disease, the progress of which was hastened by all the circumstances of
his disorderly life. The man was conscious of his degradation; he knew
how he had fallen ever since he began criminal practices; he knew the
increasing hopelessness of his resolves to have done with dangers and
recover his peace of mind. The loss of his daily work, in consequence of
irregularity, was the last thing needed to complete
{\protect\hypertarget{187}{}{}}his ruin. He did not even try to get new
employment, feeling that such a show of honest purpose was useless.
Corruption was eating to his heart; from every interview with Clem he
came away a feebler and a baser being. And upon the unresisting creature
who shared his home he had begun to expend the fury of his
self-condemnation.

He hated her because Clem bade him do so. He hated her because her
suffering rebuked him, because he must needs be at the cost of keeping
her alive, because he was bound to her.

As she moved painfully about the room he watched her with cruel,
dangerous eyes. There was a thought tormenting his brain, a terrifying
thought he had pledged himself not to dismiss, and it seemed to
exasperate him against Pennyloaf. He had horrible impulses, twitches
along his muscles; every second the restraint of keeping in one position
grew more unendurable, yet he feared to move.

Pennyloaf had the ill-luck to drop a saucer, and it broke on the floor.
In the same instant he leapt up and sprang on her, seized her brutally
by the shoulders and flung her with all his force against the nearest
wall. At her {\protect\hypertarget{188}{}{}}scream the child set up a
shrill cry, and this increased his rage. With his clenched fist he dealt
blow after blow at the half-prostrate woman, speaking no word, but
uttering a stranofe sound, such as miofht come from some infuriate
animal Penny loaf still screamed, till at length the door was thrown
open and their neighbour, Mrs, Griffin, showed herself.

``Well, I never!'' she cried, wrathfully, rushing upon Bob. ``Now you
just stop that, young man! I thought it `ud be comin to this before
long. I saw you was goin' that way.''

The mildness of her expressions was partly a personal characteristic,
partly due to Mrs. Griffin's very large experience of such scenes as
this. Indignant she might be, but the situation could not move her to
any unwonted force of utterance. Enough that Bob drew back as soon as he
was bidden, and seemed from his silence to be half-ashamed of himself.

Pennyloaf let herself lie at full length on the floor, her hands
clutched protectingly about her head; she sobbed in a quick, terrified
way, and appeared powerless to stop, even when Mrs. Griffin tried to
raise her.

``What's he been a-usin' you like this for?''

{\protect\hypertarget{189}{}{}}the woman kept asking. ``There, there
now! He shan't hit you no more, he shan't!''

Whilst she spoke Bob turned away and went from the room.

From Merlin Place he struck off into Pentonville and walked towards
King's Cross at his utmost speed. Not that he had any object in
hastening, but a frenzy goaded him along, faster, faster, till the sweat
poured from him. From King's Cross, northwards; out to Holloway, to
Hornsey. A light rain was ceaselessly falling; at one time he took off
his hat and walked some distance bareheaded, because it was a pleasure
to feel the rain trickle over him. From Hornsey by a great circuit he
made back for Islington. Here he went into a public-house, to quench the
thirst that had grown unbearable. He had but a shilling in his pocket,
and in bringing it out he was reminded of the necessity of getting more
money. He was to have met Jack Bartley to-night, long before this hour.

He took the direction for Smithfield, and soon reached the alley near
Bartholomew's Hospital where Bartley dwelt. As he entered the street he
saw a small crowd gathered about {\protect\hypertarget{190}{}{}}a
public-house door; he hurried nearer, and found that the object of
interest was a man in the clutch of two others. The latter, he perceived
at a glance, were police-officers in plain clothes; the man arrested
was---Jack Bartley himself.

Jack was beside himself with terror; he had only that moment been
brought out of the bar, and was pleading shrilly in an agony of
cowardice.

``It ain't me as made 'em! I never made one in my life! I'll tell you
who it is---I'll tell you where to find him---it's Bob Hewlett as lives
in Merlin Place! You've took the wrong man. It ain't me as made 'em!
I'll tell you the whole truth, or may I never speak another word! It's
Bob Hewett made 'em all---he lives in Merlin Place, Clerkenwell. I'll
tell you''{{------}}

Thus far had Bob heard before he recovered sufficiently from the shock
to move a limb. The officers were urging their prisoner forward,
grinning and nodding to each other, whilst several voices from the crowd
shouted abusively at the poltroon whose first instinct was to betray his
associate. Bob turned his face {\protect\hypertarget{191}{}{}}away and
walked on. He did not dare to run, yet the noises behind him kept his
heart leaping with dread. A few paces and he was out of the alley. Even
yet he durst not run. He had turned in the unlucky direction; the crowd
was still following. For five minutes he had to keep advancing, then at
last he was able to move off at right angles. The crowd passed the end
of the street.

Only then did complete panic get possession of him. With a bound forward
like that of a stricken animal he started in blind flight. He came to a
crossing, and rushed upon it regardless of the traffic. Before he could
gain the farther pavement the shaft of a cart struck him on the breast
and threw him down. The vehicle was going at a slow place, and could be
stopped almost immediately; he was not touched by the wheel A man helped
him to his feet and inquired if he were hurt.

``Hurt? No, no; it's all right.''

To the surprise of those who had witnessed the accident, he walked
quickly on, scarcely feeling any pain. But in a few minutes there came a
sense of nausea and a warm rush in his throat; he staggered against the
wall and {\protect\hypertarget{192}{}{}}vomited a quantity of blood.
Again he was surrounded by sympathising people; again he made himself
free of them and hastened on. But by now he was suffering acutely; he
could not run, so great was the pain it cost him when he began to
breathe quickly. His mouth was full of blood again.

Where could he find a hiding-place? The hunters were after him, and
however great his suffering, he must go through it in secrecy. But in
what house could he take refuge? He had not money enough to pay for a
lodging.

He looked about him; tried to collect his thoughts. By this time the
police would have visited Merlin Place; they would be waiting there to
trap him. He was tempted towards Farringdon Road Buildings; surely his
father would not betray him, and he was in such dire need of kindly
help. But it would not be safe; the police would search there.

Shooter's Gardens? There was the room where lived Pennyloaf's drunken
mother and her brother. They would not give him up. He could think of no
other refuge, at all events, and must go there if he would not drop in
the street.
