\chapter{Mad Jack's Dream}

\textsc{It} was not much more than a quarter of an hour's walk, but pain
and fear made the distance seem long; he went out of his way, too, for
the sake of avoiding places that were too well lighted. The chief
occupation of his thoughts was in conjecturing what could have led to
Bartley's arrest. Had the fellow been such a fool as to attempt passing
a bad coin when he carried others of the same kind in his pocket? Or had
the arrest of some other ``pal'' in some way thrown suspicion on Jack?
Be it as it might, the game was up. With the usual wisdom which comes
too late. Bob asked himself how he could ever have put trust in Bartley,
whom he knew to be as mean-spirited a cur as breathed. On the chance of
making things easier for himself, Jack would betray every secret in his
possession. What hope {}was there of escaping capture, even if a
hiding-place could be found for a day or two? If he had his hand on Jack
Bartley's gizzard!

Afraid to appear afraid, in dread lest his muddy clothing should attract
observation, he kept, as often as possible, the middle of the road, and
with relief saw at length the narrow archway, with its descending steps,
which was one entrance to Shooter's Gardens. As usual, two or three
loafers were hanging about here, exchanging blasphemies and filthy
vocables, but, even if they recognised him, there was not much fear of
their giving assistance to the police. With head bent he slouched past
them, unchallenged. At the bottom of the steps, where he was in all but
utter darkness, his foot slipped on garbage of some kind, and with a
groan he fell on his side.

``Let him that thinketh he standeth take heed lest he fall,'' cried a
high-pitched voice from close by.

Bob knew that the speaker was the man notorious in this locality as Mad
Jack. Raising himself with difficulty, he looked round and saw a shape
crouching in the corner.

``What is the principal thing?'' continued {}the crazy voice. ``Wisdom
is the principal thing.''

And upon that followed a long speech which to Bob sounded as gibberish,
but which was in truth tolerably good French, a language Mad Jack was
fond of using, though he never made known how he had acquired it.

Bob stumbled on, and quickly came to the house where he hoped to find a
refuge. The door was, of course, open; he went in and groped his way up
the staircase. A knock at the door of the room which he believed to be
still tenanted by Mrs. Candy and her son brought no reply. He turned the
handle, but found that the door was locked.

It was not late, only about ten o'clock. Stephen Candy could not, of
course, be back yet from his work, and the woman was probably drinking
somewhere. But he must make sure that they still lived here. Going down
to the floor below, he knocked at the room occupied by the Hope family,
and Mrs. Hope, opening the door a few inches, asked his business.

``Does Mrs. Candy still live upstairs?'' he inquired in a feigned voice,
and standing back in the darkness.

{}``For all I know.''

And the door closed sharply. He had no choice but to wait and see if
either of his acquaintances returned. For a few minutes he sat on the
staircase, but as at any moment some one might stumble over him, he went
down to the back-door, which was open, like that in front, and passed
out into the stone-paved yard. Here he seated himself on the ground,
leaning against a corner of the wall. He was suffering much from his
injury, but could at all events feel secure from the hunters.

The stones were wet, and rain fell upon him. As he looked up at the
lighted windows in the back of the house, he thought of Penny loaf, who
by this time most likely knew his danger. Would she be glad of it,
feeling herself revenged? His experience of her did not encourage him to
believe that. To all his ill-treatment she had never answered with
anything but tears and submission. He found himself wishing she were
near, to be helpful to him in his suffering.

Clem could not learn immediately what had come to pass. Finding he did
not keep his {}appointment for the day after to-morrow, she would
conclude that he had drawn back. But perhaps Jack Bartley's case would
be in the newspapers on that day, and his own name might appear in the
evidence before the magistrates; if Clem learnt the truth in that way,
she would be not a little surprised. He had never hinted to her the
means by which he had been obtaining money.

Voices began to sound from the passage within the house; several young
fellows, one or other of whom probably lived here, had entered to be out
of the rain. One voice, very loud and brutal, Bob quickly recognised; it
was that of Ned Higgs, the ruffian with whom Bartley's wife had taken
up. The conversation was very easy to overhear; it contained no
reference to the ``copping'' of Jack.

``Fag ends!'' this and that voice kept crying. Bob understood. One of
the noble company had been fortunate enough to pick up the end of a
cigar somewhere,and it was the rule among them that he who called out
``Fagends!'' established a claim for a few whiffs. In this way the
delicacy was passing from mouth to mouth. That the game should end {}in
quarrel was quite in order, and sure enough, before very long, Ned Higgs
was roaring his defiances to a companion who had seized the bit of
tobacco unjustly.

``I `ollered fag-end after Snuffy Bill!''

``You're a {{------}} liar! I did!''

``You! You're a {{------}} {{------}} {{------}}! I'll {{------}} your
{{------}} in arf a {{------}} second!''

Then came the sound of a scuffle, the thud of blows, the wild-beast
bellowing of infuriate voices. Above all could be heard the roar of Ned
Higgs. A rush, and it was plain that the combatants had gone out into
the alley to have more room. For a quarter of an hour the yells from
their drink-sodden throats echoed among the buildings. Quietness was
probably caused by the interference of police; knowing that. Bob shrank
together in his lurking-place.

When all had been still for some time he resolved to go upstairs again
and try the door, for his breathing grew more and more painful, and
there was a whirling in his head which made him fear that he might
become insensible. To rise was more difficult than he had imagined; his
head overweighted him, all but caused him to plunge forward; he groped
this {}way and that with his hands, seeking vainly for something to
cling to on the whitewashed wall. In his depth of utter misery he gave
way and sobbed several times. Then once more he had the warm taste of
blood in his mouth. Terror-stricken, he staggered into the house.

This time a voice answered to his knock. He opened the door.

The room contained no article of furniture. In one corner lay some rags,
and on the mantelpiece stood a tin teapot, two cups, and a plate. There
was no fire, but a few pieces of wood lay near the hearth, and at the
bottom of the open cupboard remained a very small supply of coals. A
candle made fast in the neck of a bottle was the source of light.

On the floor was sitting, or lying, an animated object, indescribable;
Bob knew it for Mrs. Candy. Her eyes looked up at him apprehensively.

``I want to stay the night over, if you'll let me,'' he said, when he
had closed the door.

``I've got to hide away; nobody mustn't know as I'm here.''

``You're welcome,'' the woman replied, in a voice which was horrible to
hear.

{}Then she paid no more attention to him, but leaned her head upon her
hand and began a regular moaning, as if she suffered some dull
persistent pain.

Bob crept up to the wall and let himself sink there. He could not
reflect for more than a minute or two continuously; his brain then
became a mere confused whirl. In one of the intervals of his perfect
consciousness he asked Mrs. Candy if Stephen would come here to-night.
She did not heed him till he had twice repeated the question, and then
she started and looked at him in wild fear.

``Will Stephen be coming?''

``Stephen? Yes, yes. I shouldn't wonder.'' She seemed to fall asleep as
soon as she had spoken; her head dropped heavily on the boards.

Not long after midnight the potman made his appearance. As always, on
returning from his sixteen-hour day of work, he was all but insensible
with fatigue. Entering the room, he turned his white face with an
expression of stupid wonderment to the corner in which Bob lav. The
latter raised himself to a sitting posture.

{}``That you, Bob Hewett?''

``I want to stop here over the night,'' replied the other, speaking with
difficulty. ``I can't go home. There's something up.''

``With Pennyloaf?''

``No. I've got to hide away. And I'm feeling bad---awful bad. Have you
got anything to drink?''

Stephen, having listened with the face of a somnambulist, went to the
mantelpiece and looked into the teapot. It w\^{}as empty.

``You can go to the tap in the yard,'' he said.

``I couldn't get so far. Oh, I feel bad!''

``I'll fetch you some water.''

A good-hearted animal, this poor Stephen; a very tolerable human being,
had he had fair-play. He would not abandon his wretched mother, though
to continue living with her meant hunger and cold and yet worse evils.
For himself, his life was supported chiefly on the three pints of liquor
which he was allowed every day. His arms and legs were those of a living
skeleton; his poor idiotic face was made yet more repulsive by disease.
Yet you could have seen that he was the brother of Pennyloaf; there was
Pennyloaf's submissive {}beast-of-burden look in his eyes, and his voice
had something that reminded one of hers.

``The coppers after you?'' he whispered, stooping down to Bob with the
teacup he had filled with water.

Bob nodded, then drained the cup eagerly.

``I got knocked down by a cab or something,'' he added. ``It hit me just
here. I may feel better when I've rested a bit. Haven't you got no
furniture left?''

``They took it last Saturday was a week. Took it for rent. I thought we
didn't owe nothing, but mother told me she'd paid when she hadn't. I got
leave to stop, when I showed 'em as I could pay in future; but they
wouldn't trust me to make up them three weeks. They took the furniture.
It's 'ard, I call it. I asked my guvnor if it was law for them to take
mother's bed-things, an' he said yes it was. When it's for rent they can
take everything, even to your beddin' an' tools.''

Yes; they can take everything. How foolish of Stephen Candy and his
tribe not to be born of the class of landlords! The inconvenience of
having no foothold on the earth's surface is so manifest.

{}``I couldn't say nothing to her,'' he continued, nodding towards the
prostrate woman.

``She was sorry for it, an' you can't ask no more. It was my fault for
trustin' her with the money to pay, but I get a bit careless now an
then, an' forgot. You do look bad. Bob, and there's no mistake. Would
you feel better if I lighted a bit o' fire?''

``Yes; I feel cold. I was hot just now.''

``You needn't be afraid o' the coals. Mother goes round the streets
after the coal-carts, an' you wouldn't believe what a lot she picks up
some days. You see, we're neither of us in the `ouse very often; we
don't burn much.''

He lit a fire, and Bob dragged himself near to it. In the meantime the
quietness of the house was suffering a disturbance familiar to its
denizens. Mr. Hope---you remember Mr. Hope?---had just returned from an
evening at the public-house, and was bent on sustaining his reputation
for unmatched vigour of language. He was quarrelling with his wife and
daughters; their high notes of vituperation mingled in the most
effective way with his manly thunder. To hear Mr. Hope's expressions, a
stranger would have imagined him on {}the very point of savagely
murdering all his family.

Another voice became audible. It was that of Ned Higgs, who had opened
his door to bellow curses at the disturbers of his rest.

``They'll be wakin' mother,'' said Stephen. ``There, I knew they
would.''

Mrs. Candy stirred, and after a few vain efforts to raise herself,
started up suddenly. She fixed her eyes on the fire, which was just
beginning to blaze, and uttered a dreadful cry, a shriek of mad terror.

``O God!'' groaned her son. ``I hope it ain't goin' to be one of her bad
nights. Mother, mother! what's wrong with you? See, come to the fire an'
warm yourself, mother.''

She repeated the cry two or three times, but with less violence; then,
as though exhausted, she fell face downwards, her arms folded about her
head. The moaning which Bob had heard earlier in the evening
recommenced.

Happily, it was not to be one of her bad nights. Fits of the horrors
only came upon her twice before morning. Towards one o'clock Stephen had
sunk into a sleep which scarcely any conceivable uproar could have
broken; he {}lay with his head on his right arm, his legs stretched out
at full length; his breathing was light. Bob was much later in getting
rest. As often as he slumbered for an instant, the terrible image of his
fear rose manifest before him; he saw himself in the clutch of his
hunters, just like Jack Bartley, and woke to lie quivering. Must not
that be the end of it, sooner or later? Might he not as well give
himself up to-morrow? But the thought of punishment such as his crime
receives was unendurable. It haunted him in nightmare when sheer
exhaustion had at length weighed down his eyelids.

Long before daybreak he was conscious again, tormented with thirst and
his head aching woefully. Some one had risen in the room above, and was
tramping about in heavy boots. The noise seemed to disturb Mrs. Candy;
she cried out in her sleep. In a few minutes the early riser came forth
and began to descend the stairs; he was going to his work.

A little while, and in the court below a voice shouted, ``Bill! Bill!''
Another worker being called, doubtless.

At seven o'clock Stephen roused himself. {}He took a piece of soap from
a shelf of the cupboard, threw a dirty rag over his arm, and went down
to wash at the tap in the yard. Only on returning did he address Bob.

``Feelin' any better?''

``I think so. But I'm very bad.''

``Are you goin' to stay here?''

``I don't know.''

``Got any money?''

``Yes. Ninepence. Could you get me some thin o' to drink?''

Stephen took twopence, went out, and speedily returned with a large mug
of coffee; from his pocket he brought forth a lump of cake, which had
cost a halfpenny. This, he thought, might tempt a sick appetite. His own
breakfast he would take at the coffee-shop.

``Mother'll get you anything else you want,'' he said. ``She knows
herself generally first thing in the morning. Let her take back the mug;
I had to leave threepence on it.''

So Stephen also went forth to his labour,---in his case, it may surely
be said, the curse of curses{.~.~.~.}

At this hour Pennyloaf bestirred herself {}after a night of weeping.
Last evening the police had visited her room, and Lad searched it
thoroughly. The revelation amazed her; she would not believe the charge
that was made against her husband. She became angry with Mrs. Griffin
when that practical woman said she was not at all surprised. Utterly
gone was her resentment of Bob's latest cruelty. His failure to return
home seemed to prove that he had been arrested, and she could think of
nothing but the punishment that awaited him.

``It's penal servitude,'' remarked Mrs. Griffin, frankly. ``Five, or
p'raps ten, years. I've heard of 'em get tin' sent for life.''

Pennyloaf would not believe in the possibility of this befalling her
husband. It was too cruel. There would be some pity, some mercy. She had
a confused notion of witnesses being called to give a man a good
character, and strengthened herself in the thought of what she would
say, under such circumstances, on Bob's behalf. ``He's been a good
`usband,'' she kept repeating to Mrs. Griffin, and to the other
neighbours who crowded to indulge their curiosity. ``There's nobody can
say as he ain't been a good `usband; it's a lie if they do.''

{}By eight o'clock she was at the policestation. With fear she entered
the ugly doorway and approached a policeman who stood in the ante-room.
When she had made her inquiry, the man referred her to the inspector.
She was asked many questions, but to her own received no definite reply;
she had better look in again the next morning.

``It's my belief they ain't got him,'' said Mrs. Griffin. ``He's had a
warnin' from his pals.''

Pennyloaf would dearly have liked to communicate with Jane Snowdon, but
shame prevented her. All day she stood by the house-door, looking
eagerly now this way, now that, with an unreasoning hope that Bob might
show himself. She tried to believe that he was only keeping away because
of his behaviour to her the night before; it was the first time he had
laid hand upon her, and he felt ashamed of himself. He would come back,
and this charge against him would be proved false; Pennyloaf could not
distinguish between her desire that something might happen and the
probability of its doing so.

But darkness fell upon the streets, and her {}watch was kept in rain.
She dreaded the thought of passing another night in uncertainty. Long
ago her tears had dried up; she had a parched throat and trembling,
feverish hands. Between seven and eight o'clock she went to Mrs. Griffin
and begged her to take care of the child for a little while.

``I'm goin' to see if I can hear anything about him. Somebody may know
where he is.''

And first of all she directed her steps to Shooter's Gardens. It was
very unlikely that her mother could be of any use, but she would seek
there. Afterwards she must go to Farringdon Road Buildings, though never
yet had she presented herself to Bob's father.

You remember that the Gardens had an offshoot, which was known simply as
The Court. In this blind alley there stood, throughout the day, a row of
baked-potato ovens, ten or a dozen of them, chained together, the
property of a local capitalist who let them severally to men engaged in
this business. At seven o'clock of an evening fires were wont to be
lighted under each of these baking-machines, preparatory to their being
wheeled away, each to its customary {}street-corner. Now the lighting of
fires entails the creation of smoke, and whilst these ten or twelve
ovens were getting ready to bake potatoes the Court was in a condition
not easily described. A single lamp existed for the purpose of giving
light to the alley, and at no time did this serve much more than to make
darkness visible; at present the blind man would have fared as well in
that retreat as he who had eyes, and the marvel was how those who lived
there escaped suffocation. In the Gardens themselves volumes of dense
smoke every now and then came driven along by the cold gusts; the air
had a stifling smell and a bitter taste.

Pennyloaf found nothing remarkable in this phenomenon; it is hard to say
what would have struck her as worthy of indignant comment in her world
of little ease. But near the entrance to the Court, dimly discernible
amid sagging fumes, was a cluster of people, and as everything of that
kind just now excited her apprehensions, she drew near to see what was
happening. The gathering was around Mad Jack; he looked more than
usually wild, and with one hand raised above his head was {}on the point
of relating a vision he had had the night before.

``Don't laugh! Don't any of you laugh; for as sure as I live it was an
angel stood in the room and spoke to me. There was a light such as none
of you ever saw, and the angel stood in the midst of it. And he said to
me: ` Listen, whilst I reveal to you the truth, that you may know where
you are and what you are; and this is done for a great purpose.' And I
fell down on my knees, but never a word could I have spoken. Then the
angel said: `You are passing through a state of punishment. You, and all
the poor among whom you live; all those who are in suffering of body and
darkness of mind, were once rich people, with every blessing the world
can bestow, with every opportunity of happiness in yourselves and of
making others happy. Because you made an ill use of your wealth, because
you were selfish and hard-hearted and oppressive and sinful in every
kind of indulgence,---therefore after death you received the reward of
wickedness. This life you are now leading is that of the damned; this
place to which you are confined is Hell! There is no escape for {}you.
From poor you shall become poorer; the older you grow the lower shall
you sink in want and misery; at the end there is waiting for you, one
and all, a death in abandonment and despair. This is
Hell---Hell---Hell!'''

His voice had risen in pitch, and the last cry was so terrifying that
Pennyloaf fled to be out of hearing:. She reached the house to which her
visit was, and in the dark passage leaned for a moment against the wall,
trembling all over. Then she began to ascend the stairs. At Mrs. Candy's
door she knocked gently. There was at first no answer, but when she had
knocked again, a strange voice that she did not recognise asked, ``Who's
that?'' It seemed to come from low down, as if the speaker were lying on
the floor.

``It's me,'' she replied, again trembling, she knew not with what fear.
``Mrs. Hewett---Pennyloaf.''

``Are you alone?''

She bent down, listening eagerly. ``Who's that speakin'?''

``Are you alone?''

Strange; the voice was again different, very feeble, a thick whisper.

{}``Yes, there's nobody else. Can I come in?''

There was a shuffling sound, then the keyturned in the lock. Pennyloaf
entered, and found herself in darkness. She shrank back.

``Who's there? Is it you, mother? Is it you, Stephen?''

Some one touched her, at the same time shutting the door; and the voice
whispered:

``Penny---it's me---Bob.''

She uttered a cry, stretching out her hands. A head was leaning against
her, and she bent down to lay hers against it.

``Bob! What are you doin' here? Why are you in the dark? What's the
matter, Bob?''

``I've had an accident, Penny. I feel awful bad. Your mother's gone out
to buy a candle. Have they been coming after me?''

``Yes, yes. But I didn't know you was here. I came to ask if they knew
where you was. Bob! what's happened to you `? Why are you lyin' there.
Bob?''

She had folded her arms about him, and held his face to hers, sobbing,
kissing him.

``It's all up,'' he gasped. " I've been getting worse all day. You'll
have to fetch the parish {}doctor. They'll have me, but I can't help it.
I feel as if I was going.''

``They shan't take you, Bob. Oh no, they shan't. The doctor needn't know
who you are.''

``It was a cab knocked me down, when I was running. I'm awful bad,
Peimy. You'll do something for me, won't you?''

``Oh, why didn't you send mother for me?''

The door opened. It was Mrs. Candy who entered. She slammed the door,
turned the key, and exclaimed in a low voice of alarm:

``Bob, there's the police downstairs! They come just this minute.
There's one gone to the back-door, and there's one talkin' to Mrs. Hope
at the front.''

``Then they've followed Pennyloaf,'' he replied, in a tone of despair. "
They've followed Pennyloaf.''

It was the truth. She had been watched all day, and was now tracked to
Shooter's Gardens, to this house. Mrs. Candy struck a match, and for an
instant illuminated the wretched room; she looked at the two, and they
at length saw each other's faces. Then the little flame was
extinguished, and a red spot marked the place where the remnant of the
match lay. {}``Shall I light the candle?'' the woman asked in a whisper.
Neither replied, for there was a heavy foot on the stairs. It came
nearer. A hand tried the door, then knocked loudly.

``Mrs. Candy,'' cried a stranger. The three crouched together,
terror-stricken, holding their breath. Pennyloaf pressed her husband in
an agonised embrace.

``Mrs. Candy, you're wanted on business. Open the door. If you don't
open, we shall force it.''

``No---no!'' Pennyloaf whispered in her mother's ear. ``They shan't come
in! Don't stir.''

``Are you going to open the door?''

It was a different speaker,---brief, stern. Ten seconds, and there came
a tremendous crash; the crazy door, the whole wall, quivered and cracked
and groaned. The crash was repeated, and effectually; with a sound of
ripping wood the door flew open and a light streamed into the room.

Useless, Pennyloaf, useless. That fierce kick, making ruin of your
rotten barrier, is dealt with the whole force of Law, of Society; {}you
might as well think of resisting death when your hour shall come.

``There he is,'' observed one of the men, calmly. ``Hollo! what's up?''

``You can't take him away!'' Penny loaf cried, falling down again by Bob
and clinging to him. " He's ill. You can't take him like this!''

``Ill, is he? Then the sooner our doctor sees him the better. Up you
get, my man!''

But there are some things that even Law and Society cannot command. Bob
lay insensible. Shamming? Well, no; it seemed not. Send for a stretcher,
quickly.

No great delay. Pennyloaf sat in mute anguish, Bob's head on her lap. On
the staircase was a crowd of people, talking, shouting, whistling;
presently they were cleared away by a new arrival of officials. Room for
Law and Society!

The stretcher arrived; the senseless body was carried down and laid upon
it,---a policeman at each end, and, close clinging, Pennyloaf.

Above the noise of the crowd rose a shrill, wild voice, chanting:

``All ye works of the Lord, bless ye the Lord; praise Him and magnify
Him for ever!''
