{}

{CHAPTER XI.}

JOSEPH TRANSACTS MUCH BUSINESS.

\textsc{Amid} the anguish of heart and nerve which she had to endure
whilst her grandfather lay dead in the house, Jane found and clung to
one thought of consolation. He had not closed his eyes in the bitterness
of disappointment. The end might have come on that miserable day when
her weakness threatened the defeat of all his hopes, and how could she
then have borne it? True or not, it would have seemed to her that she
had killed him; she could not have looked on his face, and all the rest
of her life would have been remorsefully shadowed. Now the dead features
were unreproachful; nay, when she overcame her childish tremors and
gazed calmly, it was easy to imagine that he smiled. Death itself had
come without pain. An old man, weary after his long journeys, {}after
his many griefs and the noble striving of his thought, surely he rested
well.

During the last days he had been more affectionate with her than was his
habit; she remembered it with gratitude. Words of endearment seldom came
to his lips, but since the reconciliation he had more than once spoken
tenderly. Doubtless he was anxious to assure her that she had again all
his confidence. Strengthening herself in that reflection, she strove to
put everything out of her mind save the duty which must henceforth
direct her. Happily, there could be no more strife with the promptings
of her weaker self; circumstances left but one path open before her; and
that, however difficult, the one she desired to tread. Henceforth memory
must dwell on one thing only in the past, her rescue by Michael Snowdon,
her nurture under his care. Though he could no longer speak, the
recollection of his words must be her unfailing impulse. In her his
spirit must survive, his benevolence still be operative.

At her wish, her father acquainted Sidney Kirkwood with what had
happened. Sidney did not visit her, but he wrote a letter, which,
{}heaving read it many times, she put carefully away to be a resource if
ever her heart failed. Mr. Percival came to the house on Monday, in the
company of Joseph Snowdon; he was sympathetic, but made no direct
reference to her position either now or in the future. Whilst he and her
father transacted matters of business in the upper rooms, Jane remained
downstairs with Mrs. Byass. Before quitting the house he asked her if
she had had any communication with Miss Lant yet.

``I ought to write and tell her,'' replied Jane.

``I will do so for you,'' said the lawyer, kindly.

And on taking leave he held her band for a moment, looking
compassionately into her pale face.

On Thursday morning there arrived a letter from Miss Lant, who happened
to be out of town and grieved that she could not return in time for the
funeral, which would be that day. There was nothing about the future,
excepting a promise that the writer would come very shortly.

Michael was buried at Abney Park {}cemetery; no ray of sunlight fell
upon his open grave, but the weather was mild, and among the budded
trees passed a breath which was the promise of spring. Joseph Snowdon
and the Byasses were Jane's only companions in the mourning-carriage;
but at the cemetery they were joined by Sidney Kirk wood. Jane saw him
and felt the pressure of his hand, but she could neither speak nor
understand anything that was said to her. On Friday morning, before she
had made a show of eating the breakfast Bessie Byass prepared for her, a
visitor arrived.

``She says her name's Mrs. Griffin,'' said Bessie, ``and she has
something very important to tell you. Do you feel you can see her?''

``Mrs. Griffin? Oh, I remember; she lives in the same house as
Pennyloaf. Yes; let her come in.''

The woman was introduced to the Byasses' parlour, which Bessie thought
more cheerful for Jane just now than the room upstairs.

``Have you heard anything of what's been goin' on with the Hewetts,
Miss?'' she began.

``No. I haven't been able to go out this week. I've had trouble at
home.''

{}``I see at once as you was in mournin', Miss, an' I'm sorry for it.
You're lookin' nothing like yourself. I don't know whether it's right to
upset you with other people's bothers, but there's that poor Mrs. Hewett
in such a state, and I said as I'd run round, `cause she seems to think
there's nobody else can come to her help as you can. I always knew as
something o' this kind `ud be `appenin'.''

``But what is it? What has happened?''

Jane felt her energies revive at this appeal for help. It was the best
thing that could have befallen, now that she was wearily despondent
after yesterday's suffering. ``Her `usband's dead, Miss.''

``Dead?''

``But that ain't the worst of it. He was took by the perlice last night,
which they wanted him for makin' bad money. I always have said as it's a
cruel thing that; `cause how can you tell who gets the bad coin, an' it
may be some pore person as can't afford to lose not a `apenny. But
that's what he's been up to, an' this long time, as it appears.''

In her dialect, which requires so many words for the narration of a
simple story, Mrs. Griffin {}told what she knew concerning Bob Hewett's
accident and capture; his death had taken place early this morning, and
Pennyloaf was all but crazy with grief. To Jane these things sounded so
extraordinary that for some time she could scarcely put a question, but
sat in dismay, listening to the woman's prolix description of all that
had come to pass since Wednesday evening. At length she called for Mrs.
Byass, for whose benefit the story was repeated. ``I'm sure you oughtn't
to go there to-day,'' was Bessie's opinion. " You've quite enough
trouble of your own, my dear.''

``And that's just what I was a-sayin', mum,'' assented Mrs. Griffin, who
had won Bessie's highest opinion by her free use of respectful forms of
address. ``I never saw no one look iller, as you may say, than the young
lady.''

``Yes, yes, I will go,'' said Jane, rising.

``My trouble's nothing to hers. Oh, I shall go at once.''

``But remember your father's coming at half-past nine,'' urged Bessie, "
and he said he wanted to speak to you particular.''

``What is the time now? A quarter to {}nine. I can be back by half-past,
I think, and then I can go again. Father wouldn't mind waiting a few
minutes. I must go at once, Mrs. Byass.''

She would hear no objection, and speedily left the house in Mrs.
Griffin's company.

At half-past nine, punctually, Mr. Snowdon's double knock sounded at the
door. Joseph looked more respectable than ever in his black frock-coat
and silk hat with the deep band. His bow to Mrs. Byass was solemn, but
gallant; he pressed her fingers like a clergyman paying a visit of
consolation, and in a subdued voice made affectionate inquiry after his
daughter.

``She has slept, I hope, poor child?''

Bessie took him into the sitting-room, and explained Jane's absence.

``A good girl; a good girl,'' he remarked, after listening with elevated
brows. " But she must be careful of her health. My visit this morning is
on matters of business; no doubt she will tell you the principal points
of our conversation afterwards. An excellent friend you have been to
her, Mrs. Byass---excellent.''

{}``I'm sure I don't see how any one could help liking her,'' said
Bessie, inwardly delighted with the expectation of hearing at length
what Jane's circumstances really were.

``Indeed, so good a friend,'' pursued Joseph, " that I'm afraid it would
distress her if she could no longer live with you. And the fact
is''---he bent forward and smiled sadly---``I'm sure I may speak freely
to you, Mrs. Byass,---but the fact is, that I'm very doubtful indeed
whether she could be happy if she lived with Mrs. Snowdon. I suppose
there's always more or less difficulty where step-children are
concerned, and in this case---well, I fear the incompatibility would be
too great. To be sure, it places me in a difficult position. Jane's very
young---very young; only just turned seventeen, poor child! Out of the
question for her to live with strangers. I had some hopes---I wonder
whether I ought to speak of it? You know Mr. Kirk wood?''

``Yes, indeed. I can't tell you how surprised I was, Mr. Snowdon. And
there seems to be such a mystery about it, too.''

Bessie positively glowed with delight in such confidential talk. It was
her dread that {}Jane's arrival might put an end to it before everything
was revealed.

``A mystery, you may well say, Mrs. Byass. I think highly of Mr. Kirk
wood, very highly; but really in this affair! It's almost too painful to
talk about---to \emph{you}''

Bessie blushed, as becomes the Englishwoman of mature years when she is
gracefully supposed to be ignorant of all it most behoves her to know.

``Well, well; he is on the point of marrying a young person with whom I
should certainly not like my daughter to associate,---fortunately there
is little chance of that. You were never acquainted with Miss Hewett?''

``Ye---yes. A long time ago.''

``Well, well; we must be charitable. You know that she is dreadfully
disfigured?''

``Disfigured? Jane didn't say a word about that. She only told me that
Mr. Kirkwood was going to marry her, and I didn't like to ask too many
questions. I hadn't even heard as she was at home.''

Joseph related to her the whole story, whilst Bessie fidgeted with
satisfaction.

``I thought,'' he added, ``that you could {}perhaps throw some light on
the mystery. We can only suppose that Kirkwood has acted from the
highest motives, but I really think---well, well, we won't talk of it
any more. I was led to this subject from speaking of this poor girl's
position. I wonder whether it will be possible for her to continue to
live in your friendly care, Mrs. Byass?''

``Oh, I shall be only too glad, Mr. Snowdon!''

``Now how kind that is of you! Of course she wouldn't want more than two
rooms.''

``Of course not.''

Joseph was going further into details, when a latch-key was heard
opening the front-door. Jane entered hurriedly. The rapid walk had
brought colour to her cheek; in her simple mourning attire she looked
very interesting, very sweet and girlish. She had been shedding tears,
and it was with unsteady voice that she excused herself for keeping her
father waiting.

``Never mind that, my dear,'' replied Joseph, as be kissed her cheek.
``You have been doing good---unselfish as always. Sit down and rest; you
must be careful not to overexert yourself.''

{}Bessie busied herself affectionately in removing Jane's hat and
jacket, then withdrew that father and child might converse in private.
Joseph looked at his daughter. His praise of her was not all mere
affectation of sentiment. He had spoken truly when he said to Scawthorne
that, but for Clem, he would ask nothing better than to settle down with
this gentle girl for his companion. Selfishness, for the most part, but
implying appreciation of her qualities. She did not love him, but he was
sincere enough with himself to admit that this was perfectly natural.
Had circumstances permitted, he would have tried hard to win some
affection from her. Poor little girl! How would it affect her when she
heard what he was going to say? He felt angry with Kirkwood; yes, truly
indignant,---men are capable of greater inconsistencies than this. She
would not have cared much about the money had Kirkwood married her; of
that he felt sure. She had lost her lover; now he was going to deprive
her of her inheritance. Cruel! Yes; but he really felt so well-disposed
to her, so determined to make her a comfortable provision for the
future; and had the {}money been hers, impossible to have regarded her
thus. Joseph was thankful to the chance which, in making him wealthy,
had also enabled him to nourish such virtuous feeling. How should he
begin? He had a bright idea, an idea worthy of him. Thrusting his hand
into his pocket, he brought out half-acrown. Then:

``Your humble friend's in a sad condition, I'm afraid, Jane?''

``She is, father.''

``Suppose you give her this! Every little helps, you know.''

Jane received the coin and murmured thanks for his kindness, but could
not help betraying some surprise. Joseph was on the watch for this. It
gave him his exquisite opportunity.

``You're surprised at me offering you money, Jane? I believe your poor
grandfather led you to suppose that---that his will was made almost
entirely in your favour?''

Jane could not reply; she searched his face.

``Would it disappoint you very much, my child,'' he continued,
sympathetically, " if it turned out that he had either altered his mind
{}or by some accident had neglected to make his will? I speak as your
father, Janey, and I think I have some knowledge of your character. I
think I know that you are as free from avarice as any one could be.''

Was it true? he began to ask himself. Why, then, had her countenance
fallen? Why did such a look of deep distress pass over it?

``The fact is, Janey,'' he continued, hardening himself a little as he
noted her expression, " your grandfather left no will. The result---the
legal result---of that is, that all his property becomes---ah---mine.
He---in fact he destroyed his will a very short time, comparatively
speaking, before he died, and he neglected to make another.
Unfortunately, you see, under these circumstances we can't be sure what
his wish was.''

She was deadly pale; there was anguish in the look with which she
regarded her father.

``I'm very sorry it pains you so, my dear,'' Joseph remarked, still more
coldly. " I didn't think you were so taken up with the thought of money.
Really, Jane, a young girl at your time of life''{{------}}

``Father, father, how can you think that? {}It wasn't to be for myself;
I thought you knew; indeed you did know!''

``But you looked so very strange, my dear. Evidently you felt''

``Yes---I feel it---I do feel it! But because it means that grandfather
couldn't get back his trust in me. Oh, it is too hard! When did he
destroy his will? When, father?''

``Ten days before his death.''

``Yes; that was when it happened. You never heard; he promised to tell
nobody. I disappointed him. I showed myself very foolish and weak
in---in something that happened then. I made grandfather think that I
was too selfish to live as he hoped,---that I couldn't do what I'd
undertaken. That was why he destroyed his will. And I thought he had
forgiven me! I thought he trusted me again! O grandfather!''

Snowdon was astonished at the explanation of his own good luck, and yet
more at Jane's display of feeling. So quiet, so reserved as he had
always known her, she seemed to have become another person. For some
moments he could only gaze at her in wonder. Never yet had he heard,
never again would he hear, {}the utterance of an emotion so profound and
so noble.

``Jane,---try and control yourself, my dear. Let's talk it over, Jane.''

``I feel as if it would break my heart. I thought I had that one thing
to comfort me. It's like losing him again---losing his confidence. To
think I should have disappointed him in just what he hoped more than
anything!''

``But you're mistaken,'' Joseph exclaimed, a generous feeling for once
getting the better of prudence. ``Listen, my dear, and I'll explain to
you. I hadn't finished when you interrupted me.''

She clasped her hands upon her lap and gazed at him in eager appeal.

``Did he say anything to you, father?''

``No,---and you may be quite sure that if he \emph{hadn't} trusted you,
he \emph{would} have said something. What's more, on the very day before
his death he wrote a letter to Mr. Percival, to say that he wanted to
make his will again. He was going to do it on the Monday,---there now!
It was only an accident; he hadn't time to do what he wished.''

{}This was making a concession which he had expressly resolved to guard
against; but Joseph's designs ripened, lost their crudity, as he saw
more and more of his daughter's disposition. He was again grateful to
her; she had made things smoother than he could have hoped.

``You really think, father, that he would have made the same will as
before?''

``Not a doubt about it, my love; not a doubt of it. In fact---now let me
set your poor little mind at rest---only two days before his
death---when was it I saw him last? Friday? Thursday?---he said to me
that he had a higher opinion of you than ever. There now, Jane!''

She would have deemed it impossible for any one to utter less than truth
in such connection as this. Her eyes gleamed with joy.

``Now you understand just how it was, Jane. What we have to talk about
now is, how we can arrange things so as to carry out your grandfather's
wish. I am your guardian, my dear. Now I'm sure you wouldn't desire to
have command of large sums of money before you are twenty-one? Just so;
your {}grandfather didn't intend it. Well, first let me ask you this
question. Would you rather live with---with your stepmother, or with
your excellent friend Mrs. Byass? I see what your answer is, and I
approve it; I fully approve it. Now suppose we arrange that you are to
have an allowance of two pounds a week? It is just possible---just
possible---that I may have to go abroad on business before long; in that
case the payment would be made to you through an agent. Do you feel it
would be satisfactory?''

Jane was thinking how much of this sum could be saved to give away.

``It seems little? But you see''{{------}}

``No, no, father. It is quite enough.''

``Good. We understand each other. Of course this is a temporary
arrangement. I must have time to think over grandfather's ideas. Why,
you are a mere child yet, Janey. Seventeen! A mere child, my dear!''

Forgetting the decorum imposed by his costume, Joseph became all but
gay, so delightfully were things arranging themselves. A hundred a year
he could very well afford just to keep his conscience at ease; and for
Jane {}it would be wealth. Excellent Mrs. Byass was as good a guardian
as could anywhere be found, and Jane's discretion forbade any fear on
her account when---business should take him away.

``Well now, we've talked quite long enough. Don't think for a moment
that you hadn't your grandfather's confidence, my dear; it would be
distressing yourself wholly without reason---wholly. Be a good
girl---why, there you see; I speak to you as if you were a child. And so
you are, poor little girl!---far too young to have worldly troubles. No,
no; I must relieve you of all that, until{{------}}. Well now, ril leave
you for to-day. Good-bye, my dear.''

He kissed her cheek, but Jane, sobbing a little, put her pure lips to
his. Joseph looked about him for an instant as if he had forgotten
something, then departed with what seemed unnecessary haste.

Jane and Mrs. Byass had a long talk before dinner-time. Mystery was at
an end between them now; they talked much of the past, more of the
future.

At two o'clock Jane received a visit from Miss Lant. This lady was
already apprised by her friend Mr. Percival of all that had come {}to
pass; she was prepared to exercise much discretion, but Jane soon showed
her that this was needless. The subject of pressing importance to the
latter was Pennyloaf 's disastrous circumstances; unable to do all she
wished, Jane was much relieved when her charitable friend proposed to
set off to Merlin Place forthwith and ascertain how help could most
effectually be given. Yes; it was good to be constrained to think of
another's sorrows.

There passed a fortnight, during which Jane spent some hours each day
with Pennyloaf. By the kindness of fate only one of Bob's children
survived him, but it was just this luckless infant whose existence made
Pennyloafs position so difficult. Alone, she could have gone back to her
slop-work, or some less miserable slavery might have been discovered;
but Pennyloaf dreaded leaving her child each day in the care of
strangers, being only too well aware what that meant. Mrs. Candy was, of
course, worse than useless; Stephen the potman had more than his work
set in looking after her. Whilst Miss Lant and Jane were straining their
wits on the hardest of all problems---to find a means of livelihood for
one {}whom society pronounced utterly superfluous, Pennyloaf most
unexpectedly solved the question by her own effort. Somewhere near the
Meat Market, one night, she encountered an acquaintance, a woman of not
much more than her own age, who had recently become a widow, and was
supporting herself (as Yell as four little ones) by keeping a stall at
which she sold second---hand children's clothing; her difficulty was to
dispose of her children whilst she was doing business at night.
Pennyloaf explained her own position, and with the result that her
acquaintance, by name Mrs. Todd, proposed a partnership. Why shouldn't
they share a room, work together with the needle in patching and making,
and by Pennyloaf's staying at home each evening keep the tribe of
youngsters out of danger? This project was carried out; the two brought
their furniture together into a garret, and it seemed probable that they
would succeed in keeping themselves alive.

But before this settlement was effected Jane's own prospects had
undergone a change of some importance. For a fortnight nothing was heard
of Joseph Snowdon in Hanover {}Street; then there came a letter from
him; it bore a Liverpool post-mark, but was headed with no address.
Joseph wrote that the business to which he had alluded was already
summoning him from England; he regretted that there had not even been
time for him to say farewell to his daughter. However, he would write to
her occasionally during his absence, and hoped to hear from her. The
allowance of two pounds a week would be duly paid by an agent, and on
receiving it each Saturday she was to forward an acknowledgment to ``Mr.
H. Jones,'' at certain reading-rooms in the City. Let her in the
meantime be a good girl, remain with her excellent friend Mrs. Byass,
and repose absolute confidence in her affectionate father---J. S.

That same morning there came also a letter from Liverpool to Mrs. Joseph
Snowdon, a letter which ran thus:

``Clem, old girl, I regret very much that affairs of pressing importance
call me away from my happy home. It is especially distressing that this
occurs just at the time when we {}were on the point of taking our house,
in which we hoped to spend the rest of our lives in bliss. Alas, that is
not to be! Do not repine, and do not break the furniture in the
lodgings, as your means will henceforth be limited, I fear. You will
remember that I was in your debt, with reference to a little affair
which happened in Clerkenwell Close, not such a long time ago; please
accept this intimation as payment in full. When I am established in the
country to which business summons me, I shall of course send for you
immediately, but it may happen that some little time will intervene
before I am able to take that delightful step. In the meanwhile your
mother will supply you with all the money you need; she has full
authority from me to do so. All blessings upon you, and may you be
happy.---With tears I sign myself,

\textsc{``Your broken-hearted Husband.''}{⁠}

Joseph's absence through the night had all but prepared Clem for
something of this kind, yet he had managed things so well that up to the
time of his departure she had not been able to remark a single
suspicious circumstance {}unless, indeed, it were the joyous
affectionateness with which he continued to behave. She herself had been
passing through a time of excitement and even of suffering. When she
learned from the newspaper what fate had befallen Bob Hewett, it was as
though some one had dealt her a half-stunning blow; in her fierce animal
way she was attached to Bob, and for the first time in her life she knew
a genuine grief. The event seemed at first impossible; she sped hither
and thither, making inquiries, and raged in her heart against every one
who confirmed the newspaper report. Combined with the pain of loss was
her disappointment at the frustration of the scheme Bob had undertaken
in concert with her. Brooding on her deadly purpose, she had come to
regard it as a certain thing that before long her husband would be
killed. The details were arranged; all her cunning had gone to the
contrivance of a plot for disguising the facts of his murder. Savagely
she had exulted in the prospect, not only of getting rid of him, but of
beins revensfed for her old humiliation. A thousand times she imaofined
herself in Bob's lurking-place, raising the weapon, {}striking the
murderous blow, rifling the man's pockets to mislead those who found his
body, and had laughed to herself triumphantly. Joseph out of the way,
the next thing was to remove Pennyloaf. Oh, that would easily have been
contrived. Then she and Bob would have been married.

Very long since Clem had shed tears, but she did so this day when there
was no longer a possibility of doubting that Bob was dead. She shut
herself in her room and moaned like a wild beast in pain. Joseph could
not but observe, when he came home, that she was suffering in some
extraordinary way. When he spoke jestingly about it, she all but rushed
upon him with her fists. And in the same moment she determined that he
should not escape, even if she had to murder him with her own hands.
From that day her constant occupation was searching the newspapers to
get hints about poisons. Doubtless it was as well for Joseph to be
speedy in his preparations for departure.

She was present in the police-court when Jack Bartley came forward to be
dealt with. Against him she stored up hatred and the {}resolve of
vengeance; if it were years before she had the opportunity, Jack should
in the end pay for what he had done.

And now Joseph had played her the trick she anticipated; he had saved
himself out of her clutches, and had carried off all his money with him.
She knew well enough what was meant by his saying that her mother would
supply what she needed; very likely that he had made any such
arrangement! You should have heard the sterling vernacular in which Clem
gave utterance to her feelings as soon as she had deciphered the mocking
letter!

Without a minute's delay she dressed and left the house. Having a few
shillings in her pocket, she took a cab at King's Cross and bade the
driver drive his hardest to Clerkenwell Close. Up Pentonville Hill
panted the bony horse, Clem swearing all the time because it could go no
quicker. But the top was reached; she shouted to the man to whip, whip!
By the time they pulled up at Mrs. Peckover's house Clem herself
perspired as profusely as the animal.

Mrs. Peckover was at breakfast, alone.

``Read that, will you? Read that!'' roared {}Clem, rushing upon her and
dashing the letter in her face.

``Why, you mad cat!'' cried her mother, starting up in anger. ``What's
wrong with you now?''

``Read that there letter! That's \emph{your} doin', that is! Read it!
Read it!''

Half-frightened, Mrs. Peckover drew away from the table and managed to
peruse Joseph's writing. Having come to the end, she burst into jeering
laughter.

``He's done it, has he? He's took his `ook, has he? \emph{What} did I
tell you? Don't swear at me, or I'll give you something to swear
about---such languidge in a respectable `ouse! Ha, ha! What did I tell
you'? You wouldn't take \emph{my} way. Oh no, you must go off and be
independent. \emph{Serve} you right! Ha, ha! Serve you right I You'll
get no pity from me.''

``You 'old your jaw, mother, or I'll precious soon set my marks on your
ugly old face! What does he say there about you 1 You're to pay me
money. He's made arrangements with you. Don't try to cheat me, or
I'll---soon have a summons out against you. The {}letter's proof; it's
lawyer's proof. You try to cheat me and see.''

Clem had sufficient command of her faculties to devise this line of
action. She half believed, too, that the letter would be of some legal
efficacy, as against her mother.

``You bloomin' fool! '' screamed Mrs. Peckover. ``Do you think I was
born yesterday? Not one farden do you get out of me if you starve in the
street---not one farden! It's my turn now. I've had about enough o' your
cheek an' your hinsults. You'll go and work for your livin', you great
cart-horse!''

``Work! No fear! I'll set the perlice after him.''

``The perlice! What can they do?''

``Is it law as he can go off and leave me with nothing to live on?''

``Course it is! Unless you go to the work'us an' throw yourself on the
parish. Do, do! Oh my! shouldn't I like to see you brought down to the
work'us, like Mrs. Igginbottom, the wife of the cat's-meat man, him as
they stuck up wanted for desertion!''

``You're a liar!'' Clem shouted. ``I can make \emph{you} support me
before it comes to that.''

{}The wrangle continued for some time longer; then Clem bethought
herself of another person with whom she must have the satisfaction of
speaking her mind. On the impulse, she rushed away, out of Clerkenwell
Close, up St. John Street Road, across City Road, down to Hanover
Street, literally running for most of the time. Her knock at Mrs.
Byass's door was terrific.

``I want to see Jane Snowdon,'' was her address to Bessie.

``Do you? I think you might have knocked more like civilisation,''
replied Mrs. Byass, proud of expressing herself with superior
refinement.

But Clem pushed her way forward. Jane, alarmed at the noise, showed
herself on the stairs.

``You just come `ere!'' cried Clem to her.

``I've got something to say to \emph{you}, Miss!''

Jane was of a sudden possessed with terror, the old terror with which
Clem had inspired her years ago. She shrank back, but Bessie Byass was
by no means disposed to allow this kind of thing to go on in her house.

``Mrs. Snowdon,'' she exclaimed, ``I don't {}know what your business may
be, but if you can't behave yourself, you'll please to go away a bit
quicker than you came. The idea! Did any one ever hear!''

``I shan't go till I choose,'' replied Clem, ``and that won't be till
I've had my say with that little{{------}}! Where's your father, Jane
Snowdon? You just tell me that.''

``My father,'' faltered Jane, in the silence.

``I haven't seen him for a fortnight.''

``You haven't, eh? Little liar! It's what I used to call you when you
scrubbed our kitchen floor, and it's what I call you now. D'you remember
when you did the `ouse-work, `an slept under the kitchen table? D'you
remember, eh? Haven't seen him for a fortnight, ain't you? Oh, he's a
nice man, is your father! He ran away an' deserted your mother. But he's
done it once too often. \emph{I}'ll precious soon have the perlice after
him! Has he left you to look after yourself? Has he, eh? You just tell
me that!''

Jane and Mrs. Byass stared at each other in dismay. The letter that had
come this morning enabled them to guess the meaning of Clem's fury. The
latter interpreted their {}looks as an admission that Jane too was a
victim. She laughed aloud.

``How does it taste, little liar, eh? A second disappointment! You
thought you was a-goin to have all the money; now you've got none, and
you may go back to Whitehead's. They'll be glad to see you, will
Whitehead's. Oh, he's a nice man, your father! Would you like to know
what's been goin' on ever since he found out your old grandfather? Would
you like to know how he put himself out to prevent you an' that Kirkwood
feller gettin' married, just so that the money mightn't get into other
people's `ands? Would you like to know how my beast of a mother and him
put their `eds together to see how they could get hold of the bloomin
money? An' \emph{you} thought you was sure of it, didn't you? Will you
come with me to the perlice-station, just to help to describe what he
looks like? An affectionate father, ain't he? Almost as good as he is a
`usband. You just listen to me, Jane Snowdon. If I find out as you're
havin' money from him, I'll be revenged on you, mind that! I'll be
revenged on you! D'you remember what my hand feels {}like? You've had it
on the side of your {{------}} 'ed often enough. You just look out for
yourself!''

``And you just turn out of my house,'' cried Bessie, scarlet with wrath.
``This minute! Sarah! Sarah! Run out by the arey-steps and fetch a
p'liceman, this minute! The idea!''

Clem had said her say, however, and with a few more volleys of atrocious
language was content to retire. Having slammed the door upon her, Bessie
cried in a trembling voice:

``Oh, if only Sam had been here! My, how I should have liked Sam to have
been here! \emph{Wouldn't} he have given her something for herself! Why,
such a creature oughtn't be left loose. Oh, if Sam had been here!''

Jane had sat down on the stairs; her face was hidden in her hands. That
brutal voice had carried her back to her wretched childhood; everything
about her in the present was unreal in comparison with the terrors, the
hardships, the humiliations revived by memory. As she sat at this
moment, so had she sat many a time on the cellar-steps at Mrs.
Peckover's. So powerfully was her imagination affected that she had a
feeling as {}if her hands were grimy from toil, as if her limbs ached.
Oh that dreadful voice! Was she never, never to escape beyond hearing of
it?

``Jane, my dear, come into the sittingroom,'' said Bessie. ``No wonder
it's upset you. What \emph{can} it all mean?''

The meaning was not far to seek; Jane understood everything,---yes, even
her father's hypocrisies. She listened for a few minutes to her friend's
indignant exclamations, then looked up, her resolve taken.

``Mrs. By ass, I shall take no more money. I shall go to work again and
earn my living. How thankful I am that I can!''

``Why, what nonsense are you talking, child! Just because that---that
\emph{creature}{{------}}Why, I've no patience with you, Jane! As if she
durst touch you! Touch you? I'd like to see her indeed.''

``It isn't that, Mrs. Byass. I can't take money from father. I haven't
felt easy in my mind ever since he told me about it, and now I
\emph{can't} take the money. Whether it's true or not, all she said, I
should never have a night's rest if I consented to live in this way.''

{}``Oh, you \emph{don't} really mean it, Jane?''

Bessie all but sobbed with vexation.

``I mean it, and I shall never alter my mind. I shall send back the
money, and write to the man that he needn't send any more. However often
it comes, I shall always return it. I couldn't, I couldn't live on that
money I Never ask me to, Mrs. Byass.''

Practical Bessie had already begun to ask herself what arrangement Jane
proposed to make about lodgings. She was no Mrs. Peckover, but neither
did circumstances allow her to disregard the question of rent. It cut
her to the heart to think of refusing an income of two pounds per week.

Jane too saw all the requirements of the case. ``Mrs. Byass, will you
let me have one room,---my old room upstairs? I have been very happy
there, and I should like to stay if I can. You know what I can earn; can
you afford to let me live there? I'd do my utmost to help you in the
house; I'll be as good as a servant, if you can't keep Sarah. I should
so like to stay with you!''

"You just let me hear you talk about {}leaving, that's all! Wait till
I've talked it over with Sam."

Jane went upstairs, and for the rest of the day the house was very
quiet.

Not Whitehead's; there were other places where work might he found. And
before many days she had found it. Happily there were no luxuries to be
laid aside; her ordinary dress was not too good for the workroom. She
had no habits of idleness to overcome, and an hour at the table made her
as expert with her fingers as ever.

Returning from the first day's work, she sat in her room---the little
room which used to be hers---to rest and think for a moment before going
down to Bessie's supper-table. And her thought was:

``He too is just coming home from work. Why should my life be easier
than his?''
