\hypertarget{headerContainer}{}
\hypertarget{navigationHeader}{}
\protect\hypertarget{headerprevious}{}{←\href{/wiki/The_Nether_World/Volume_3/Chapter_5}{Volume
3, Chapter 5}}

\textbf{\protect\hypertarget{header_title_text}{}{\href{/wiki/The_Nether_World}{The
Nether World}}} \emph{by
\href{/wiki/Author:George_Gissing}{\protect\hypertarget{header_author_text}{}{{George
Gissing}}}}\\
\protect\hypertarget{header_section_text}{}{Volume 3, Chapter 6}

\protect\hypertarget{headernext}{}{\href{/wiki/The_Nether_World/Volume_3/Chapter_7}{Volume
3, Chapter 7}→}

\hypertarget{navigationNotes}{}

\hypertarget{ws-data}{}
\protect\hypertarget{ws-article-id}{}{2204009}\protect\hypertarget{ws-title}{}{\href{/wiki/The_Nether_World}{The
Nether World} --- \emph{Volume 3, Chapter
6}}\protect\hypertarget{ws-author}{}{George Gissing}

{\protect\hypertarget{120}{}{}}

{CHAPTER VI.}

A FALL FROM THE IDEAL.

\textsc{Joseph Snowdon} waxed daily in respectability. He was, for one
thing, clothing himself in flesh, and, though still anything but a
portly man, bore himself as becomes one who can indulge a taste for
eating and drinking; his step was more deliberate, he no longer
presented the suppleness of limb that so often accompanies a needy
condition in the man of wits, he grew attentive to his personal
equipment, he was always well combed and well shaven, and generally, in
hours of leisure, you perceived a fragrance breathing from his
handkerchief. Nor was this refinement addressed only to the public. To
Clem he behaved with a correctness which kept that lady in a state of
acute suspicion; not seldom he brought her a trifling gift, which he
would offer with compliments, and he made a point of consulting her
{\protect\hypertarget{121}{}{}}pleasure or convenience in all matters
that affected them in common. A similar dignity of bearing marked his
relations with Hanover Street. When he entered Jane's parlour it was
with a beautiful blending of familiarity and courtesy; he took his
daughter's hand with an air of graceful affection, retaining it for a
moment between his own, and regarding: her with a gentle smile which
hinted the pride of a parent. In speaking with the old man he habitually
subdued his voice, respectfully bending forward, solicitously watching
the opportunity of a service. Michael had pleasure in his company and
conversation. Without overdoing it, Joseph accustomed himself to speak
of philanthropic interests. He propounded a scheme for supplying the
poor with a certain excellent filter at a price all but nominal; who did
not know the benefit to humble homes of pure water for use as a
beverage? The filter was not made yet, but Lake, Snowdon, \& Co. had it
under their consideration.

Michael kept his room a good deal in these wretched days of winter, so
that Joseph had no difficulty in obtaining private interview's with his
daughter. Every such occasion he {\protect\hypertarget{122}{}{}}used
assiduously, his great end being to possess himself of Jane's
confidence. He did not succeed quite so well with the girl as with her
grandfather; there was always a reserve in her behaviour which as yet he
found it impossible to overcome. Observation led him to conclude that
much of this arose from the view she took of his relations with Sidney
Kirkwood. Jane was in love with Sidney; on that point he could have no
doubt; and in all likelihood she regarded him as unfriendly to Sidney's
suit,---women are so shrewd in these affairs. Accordingly, Joseph made
it his business by artful degrees to remove this prepossession from her
mind. In the course of this endeavour he naturally pressed into his
service the gradually discovered fact that Sidney had scruples of
conscience regarding Jane's fortune. Marvellous as it appeared to him,
he had all but come to the conclusion that this \emph{was} a fact. Now,
given Jane's character, which he believed he had sounded; given her love
for Kirkwood, which was obviously causing her anxiety and unhappiness;
Joseph saw his way to an admirable piece of strategy. What could be
easier, if he played his cards well and patiently enough,
{\protect\hypertarget{123}{}{}}than to lead Jane to regard the fortune
as her most threatening enemy? Valuable results might come of that,
whether before or after the death of the old man.

The conversation in which he first ventured to strike this note
undisguisedly took place on the same evening as that unpleasant scene
when Sidney as good as quarrelled with him,---the evening before the day
on which Sidney asked Clara Hewett to be his wife. Having found Jane
alone, he began to talk in his most paternal manner, his chair very near
hers, his eyes fixed on her sewing. And presently, when the ground was
prepared:

``Jane, there's something I've been wanting to say to you for a long
time. My dear, I'm uneasy about you.''

``Uneasy, father?'' and she glanced at him nervously.

``Yes, I'm uneasy. But whether I ought to tell you why, Tm sure I don't
know. You're my own child, Janey, and you become dearer to me every day;
but---it's hard to say it---there naturally isn't all the confidence
between us that there might have been if---well, well, I won't speak of
that.''

{\protect\hypertarget{124}{}{}}``But won't you tell me what makes you
anxious?''

He laid the tips of his fingers on her head.

``Janey, shall you be offended if I speak about Mr. Kirkwood?''

``No, father.''

She tried in vain to continue sewing.

``My dear,---I believe there's no actual enogagement between you?''

``Oh no, father,'' she replied, faintly.

``And yet---don't be angry with me, my child---I think you are something
more than friends?''

She made no answer.

``And I can't help thinking, Janey,---I think about you very often
indeed,---that Mr. Kirkwood has rather exaggerated views about the
necessity of---of altering things between you.''

Quite recently Joseph had become aware of the understanding between
Michael and Kirkwood. The old man still hesitated to break the news to
Jane, saying to himself that it was better for Sidney to prepare her by
the change in his behaviour.

``Of altering things?'' Jane repeated, under her breath.

{\protect\hypertarget{125}{}{}}``It seems to me wrong---wrong to both of
you,'' Joseph pursued, in a pathetic voice. ``I can't help noticing my
child's looks. I know she isn't what she used to be, poor little girl!
And I know Kirkwood isn't what he used to be. It's very hard, and I feel
for you---for both of you.''

Jane sat motionless, not daring to lift her eyes, scarcely daring to
breathe.

``Janey.''

``Yes, father.''

``I wonder whether I'm doing wrong to your grandfather in speaking to
you confidentially like this? I can't believe he notices things as I do;
he'd never wish you to be unhappy.''

``But I don't quite understand, father. What do you mean about Mr.
Kirkwood? Why should he''{{------}}

The impulse failed her. A fear which she had harboured for many weary
days was being confirmed, and she could not ask directly for the word
that would kill hope.

``Have I a right to tell you? I thought perhaps you understood.''

``As you have gone so far, I think you must explain. I don't see how you
can be doing wrong.''

{\protect\hypertarget{126}{}{}}``Poor Kirkwood! You see, he's in such a
delicate position, my dear. I think myself that he's acting rather
strangely, after everything; but it's---it's your money, Jane. He
doesn't think he ought to ask you to marry him, under the
circumstances.''

She trembled.

``Now who should stand by you, in a case like this, if not your own
father? Of course he can't say a word to you himself; and of course you
can't say a word to him; and altogether it's a pitiful business.''

Jane shrank from discussing such a topic with her father. Her next words
were uttered with difficulty.

``But the money isn't my own---it'll never be my own. He---Mr. Kirkwood
knows that.''

``He does, to be sure. But it makes no difference. He has told your
grandfather, my love, that---that the responsibility would be too great.
He has told him distinctly that everything's at an end---everything that
\emph{might} have happened.''

She just looked at him, then dropped her eyes on her sewing.

``Now, as your father, Janey, I know it's
{\protect\hypertarget{127}{}{}}right that you should be told of this. I
feel you're being very cruelly treated, my child. And I wish to goodness
I could only see any way out of it for you both. Of course I'm powerless
either for acting or speaking; you can understand that. But I want you
to think of me as your truest friend, my love.''

More still he said, but Jane had no ears for it. When he left her, she
bade him good-bye mechanically, and stood on the same spot by the door,
without thought, stunned by what she had learnt.

That Sidney would be impelled to such a decision as this she had never
imagined. His reserve whilst yet she was in ignorance of her true
position she could understand; also his delaying for a while even after
everything had been explained to her. But that he should draw away from
her altogether seemed inexplicable, for it implied a change in him which
nothing had prepared her to think possible. Unaltered in his love, he
refused to shore the task of her life, to aid in the work which he
regarded with such fervent sympathy. Her mind was not subtle enough to
conceive those objections to Michael's idea
{\protect\hypertarget{128}{}{}}which had weighed with Sidney almost from
the first, for though she had herself shrunk from the great undertaking,
it was merely in weakness,---a reason she never dreamt of attributing to
him. Nor had she caught as much as a glimpse of those base, scheming
interests, contact with which had aroused Sidney's vehement disgust. Was
her father to be trusted? This was the first question that shaped itself
in her mind. He did not like Sidney; that she had felt all along, as
well as the reciprocal coldness on Sidney's part. But did his
unfriendliness go so far as to prompt him to intervene with untruths?
``Of course you can't say a word to him''---that remark would bear an
evil interpretation, which her tormented mind did not fail to suggest.
Moreover, he had seemed so anxious that she should not broach the
subject with her grandfather. But what constrained her to silence? If,
indeed, he had nothing but her happiness at heart, he could not take it
ill that she should seek to understand the whole truth, and Michael must
tell her whether Sidney had indeed thus spoken to him.

Before she had obtained any show of control over her agitation Michael
came into the room. {\protect\hypertarget{129}{}{}}Evening was the old
man's best time, and when he had kept his own chamber through the day he
liked to come and sit with Jane as she had her supper.

``Didn't I hear your father's voice?'' he asked, as he moved slowly to
his accustomed chair.

``Yes. He couldn't stay.''

Jane stood in an attitude of indecision. Having seated himself, Michael
glanced at her. His regard had not its old directness; it seemed
apprehensive, as if seeking to probe her thought.

``Has Miss Lant sent you the book she promised?''

``Yes, grandfather.''

This was a recently published volume dealing with charitable enterprise
in some part of London. Michael noticed with surprise the uninterested
tone of Jane's reply. Again he looked at her, and more searchingiy.

``Would you like to read me a little of it?''

She reached the book from a side-table, drew near, and stood turning the
pages. The confusion of her mind was such that she could not have read a
word with understanding. Then she spoke, involuntarily.

{\protect\hypertarget{130}{}{}}``Grandfather, has Mr. Kirkwood said
anything more---about me?''

The words made painful discord in her ears, but instead of showing
heightened colour she grew pallid. Holding the book partly open, she
felt all her nerves and muscles strained as if in some physical effort;
her feet were rooted to the spot.

``Have you heard anything from him?'' returned the old man, resting his
hands on the sides of the easy-chair.

``Father has been speaking about him. He says Mr. Kirk wood has told you
something.''

``Yes. Come and sit down by me, Jane.''

She could not move nearer. Though unable to form a distinct conception,
she felt a foreboding of what must come to pass. The dread failure of
strength was more than threatening her; her heart was sinking, and by no
effort of will could she summon the thoughts that should aid her against
herself.

``What has your father told you?'' Michael asked, when he perceived her
distress. He spoke with a revival of energy, clearly, commandingly.

``He says that Mr. Kirkwood wishes you to
{\protect\hypertarget{131}{}{}}forget what he told you, and what you
repeated to me.''

``Did he give you any reason? ''

``Yes. I don't understand, though.''

``Come here by me, Jane. Let's talk about it quietly. Sidney doesn't
feel able to help you as he thought he could. We mustn't blame him for
that; he must judge for himself He thinks it'll be better if you
continue to be only friends.''

Jane averted her face, his steady look being more than she could bear.
For an instant a sense of uttermost shame thrilled through her, and
without knowing what she did, she moved a little and laid the book down.

``Come here, my child,'' he repeated, in a gentler voice.

She approached him.

``You feel it hard. But when you've thought about it a little you won't
grieve; I'm sure you won't. Remember, your life is not to be like that
of ordinary women. You've higher objects before you, and you'll find a
higher reward. You know that, don't you? There's no need for me to
remind you of what we've talked about so often, is there? If it's
{\protect\hypertarget{132}{}{}}a sacrifice, you're strong enough to face
it; yes, yes, strong enough to face more than this, my Jane is! Only fix
your thoughts on the work you're going to do. It'll take up all your
life, Jane, won't it? You'll have no time to give to such things as
occupy other women,---no mind for them.''

His grey eyes searched her countenance with that horrible intensity of
fanaticism which is so like the look of cruelty, of greed, of any
passion originating in the baser self Unlike too, of course, but it is
the pitilessness common to both extremes that shows most strongly in an
old, wrinkled visage. He had laid his hand upon her. Every word was a
stab in the girl's heart, and so dreadful became her torture, so
intolerable the sense of being drawn by a fierce will away from all she
desired, that at length a cry escaped her lips. She fell on her knees by
him, and pleaded in a choking voice.

``I can't! Grandfather, don't ask it of me! Give it all to some one
else---to some one else! I'm not strong enough to make such a sacrifice.
Let me be as I was before!''

Michael's face darkened. He drew his hand away and rose from the seat;
with more than {\protect\hypertarget{133}{}{}}surprise, with anger and
even bitterness, he looked down at the crouching girL She did not sob;
her face buried in her arms, she lay against the chair, quivering,
silent.

``Jane, stand up and speak to me!''

She did not move.

``Jane!''

He laid his hand on her. Jane raised her head, and endeavoured to obey
him; in the act she moaned and fell insensible. Michael strode to the
door and called twice or{]} thrice for Mrs. By ass; then he stooped by
the lifeless girl and supported her head. Bessie was immediately at
hand, with a cry of consternation, but also with helpful activity.

``Why, I thought she'd got over this; it's a long time since she was
took last, isn't it? Sam's downstairs, Mr. Snowdon; do just shout out to
him to go for some brandy. Tell him to bring my smelling-bottle first,
if he knows where it is,---I'm blest if I do! Poor thing! She ain't been
at all w\^{}ell lately, and that's the truth.''

The truth, beyond a doubt. Pale face, showing now the thinness which it
had not wholly outgrown, the inheritance from
{\protect\hypertarget{134}{}{}}miserable childhood; no face of a stern
heroine, counting as idle all the natural longings of the heart,
consecrated to a lifelong combat with giant wrongs. Nothing better nor
worse than the face of one who can love and must be loved in turn.

She came to herself, and at the same moment Michael went from the room.

``There now; there now,'' crooned Bessie, with much patting of the hands
and stroking of the cheeks. ``Why, what's come to you, Jane? Cry away;
don't try to prevent yourself; it'll do you good to cry a bit. Of
course, here comes Sam with all sorts of things, when there's no need of
him. He's always either too soon or too late, is Sam. Just look at him,
Jane; now if \emph{he} don't make you laugh, nothing will!''

Mr. Byass retired, shamefaced. Leaning against Bessie's shoulder, Jane
sobbed for a long time, sobbed in the misery of shame. She saw that her
grandfather had gone away. How should she ever face him after this? It
was precious comfort to feel Bessie's sturdy arms about her, and to hear
the foolish affectionate words, which asked nothing but that
{\protect\hypertarget{135}{}{}}she should take them kindly and have done
with her trouble.

``Did grandfather tell you how it was?'' she asked, with a sudden fear
lest Bessie should have learnt her pitiful weakness.

``Why, no; how did it come?''

``I don't know. We were talking. I can stand up now, Mrs. Byass, thank
you. I'll go up to my room. I've forgotten the time; is it late?''

It was only nine o'clock. Bessie would have gone upstairs with her, but
Jane insisted that she was quite herself. On the stairs she trod as
lightly as possible, and she closed her door without a sound. Alone, she
again gave way to tears. Michael's face was angry in her memory; he had
never looked at her in that way before, and now he would never look with
the old kindness. What a change had been wrought in these few minutes!

And Sidney never anything but her friend,---cold, meaningless word! If
he knew how she had fallen, would that be likely to bring him nearer to
her? She had lost both things, that was all.
